### 2011-11-13 changes.tex
Submitted to TPAMI

\documentclass{article}
\begin{document}
\title{Summary of changes}
\maketitle
The manuscript for this submission is based on a paper that
was presented at the 2011 IEEE Workshop on Statistical Signal
Processing (June 28 - June 30, 2011) in Nice, France. The original
paper that was presented at the workshop is of 4 double column pages
in length and contains a brief discussion of the methodology along
with an example for inference on time series of graphs.

The main differences between the current manuscript and the paper from the
workshop are:
\begin{itemize}
\item A more comprehensive introduction and discussion of
  methodology. We describe the underlying model in more details.
\item Theoretical results and their derivations are now included. They
  account for most of the changes.
\item The experimental results are further refined. The main running example
  is expanded to include power approximations for number of triangles.
\end{itemize}
\end{document}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: