### 2012-06-22 decision.tex

06-Apr-2012

Prof. Carey Priebe
Johns Hopkins University
Whitehead Hall
Baltimore
Maryland
United States
21218-2682

Manuscript: T-SP-13180-2012 "Attribute fusion in a latent process
model for time series of graphs"

Dear Prof. Carey Priebe,

Your manuscript, T-SP-13180-2012 "Attribute fusion in a latent process
model for time series of graphs", which you submitted to the
Transactions on Signal Processing has been reviewed. The comments from
the reviewers are included at the bottom of this letter (**See note
below about attachments).

In view of the criticisms of the reviewers, I must reject the
manuscript for publication in the IEEE Transactions on Signal
Processing at this time. However, a new manuscript may be submitted
which takes into consideration these comments.

Your new manuscript must be submitted back to Manuscript Central
http://mc.manuscriptcentral.com/tsp-ieee together with a
point-by-point reply that explains how you addressed the reviewers'
comments.

Please note that resubmitting your manuscript does not guarantee
eventual acceptance, and that your resubmission will be subject to
re-review by the reviewers before a decision is rendered.

You will be unable to make your revisions on the originally submitted
version of your manuscript. Once you have revised your manuscript, log
into http://mc.manuscriptcentral.com/tsp-ieee and enter your Author
Center, where you will find your manuscript title listed under
"Manuscripts with Decisions". Under "Actions", click on "Create a
Resubmission" then follow the steps for resubmitting your manuscript.

If you have any questions, please contact the Admin for this journal
Ziggy Kowalski at z.kowalski@ieee.org.

Sincerely,

Prof. Maciej Niedzwiecki
Associate Editor
IEEE Transactions on Signal Processing
maciekn@eti.pg.gda.pl, maciekn@pg.gda.pl

Ziggy Kowalski
Coordinator Society Publications
IEEE Signal Processing Society
z.kowalski@ieee.org

** {This applies to SUBMITTING AUTHOR Accounts ONLY: You can find any
   possible ATTACHMENTS FROM THE REVIEWERS by going to the "Manuscript
   with Decisions" status link in your Author Center and clicking on
   "view decision letter". They are located at the bottom of decision
   letter under "Files attached" heading}

Reviewer Comments:

Reviewer: 1

Recommendation: R - Reject (Paper Is Not Of Sufficient Quality Or
Novelty To Be Published In This Transactions)

Comments:
Let's say that an graph edge can carry one of k different
attributes. Then each vertex is equipped with a vector of k elements:
there is an entry in this vector representing the affinity of the
vertex to act as an endpoint for each of the k possible types of
attributed edges. Based on these hidden vectors the probability of two
vertices to be connected is the dot product of their independent
vectors and more specifically the conditional probability that a
connecting edge will be labelled by attribute i is simply the product
of the ith vector entries. Practically, given the vectors attached to
each vertex, we can then compute "wiring" and "labelling"
probabilities and thus construct a graph instance based on them.

For a time series of graphs the authors extend to hidden vectors with
time-varying entries. Their goal is to identify the time when a
chatter anomaly - defined as perturbations in the links connecting
only a subset of the vertices - occurs in this series. At this point
they introduce extra assumptions in how the vectors can vary in time:
In any time interval $[t, t+1)$, a stationary, continuous-time Markov
process with k+1 states is assumed behind the scenes for computing the
time evolution of the vector (the extra state is absence of connection
for the vertex). Identical transition-matrix ($Q_1$) Markov processes
are suggested for all nodes before the anomaly, while after the
anomaly occurs they assume two Markov processes over the graph: the
previous one with stationary distribution $\pi_1$ running at nodes
with no incident link perturbations and $\pi_2$ at the others.

They next introduce two approximations of this assumed model: The
first order approximation considers attaching a hidden vector equal to
the stationary distribution of the Markov chain at works there; this
means that only perturbation-incident vertices will shift from $\pi_1$
to $\pi_2$ vectors at the time the anomaly happens. The second order
approximation assumes hidden vectors sampled from within an interval
around the first order approximation (stationary-distribution)
vectors.

The next step is to define their test statistics. For a given graph
they count the number of edges, the number of triangles, the maximum
degree and the maximum number of edges between neighbors over all
vertices, however inserting a weighting vector $\lambda$ with k
entries as a parameter, effectively changing the contribution of
different-attribute edges in the counting process. For each of these
four quantities they define a respective test statistic, as the
normalized deviation of the quantity from its average over a period of
the last $l$ graph history snapshots.

Selecting the weighting vector $\lambda$ that maximizes the power of
their approach is their next stop. This is the probability their test
will reject the null hypothesis when the null hypothesis is actually
false (the null hyposthesis states that the anomaly occurs after the
actual time it occurs, and this is obviously erroneous). They compute
this maximum-power $\lambda$ by actually finding the distributions
followed by their test statistics (Student t for the first two and
Gumbel for the rest two - maximum related - ones).

In the experimental section they work with two graphs, with 2
attributes each (k=2). A synthetic one where they explicitly set the
two Markov processes and then explore the dependence of the power for
exponentially decreasing waiting times for the processes (increasing
r) for their exact model and its two approximations. They also reduce
the Enron dataset into a k=2 attributed graph and plot their test
statistics as a function of $\theta$ (the magnitude of optimum
$\lambda$ vector is known to be 1, so for k=2 dimensions it could be
parameterized as $(\cos(\theta), \sin(\theta))$.

Although the idea of linearly fusing attributes seems appealing and
the authors present their ideas in a technically sound way, nicely
structured and well supported by theoretical analysis and derivations,
I have some critical reservations:

1. The practicality of the approach: Both examples consider the case
of graphs with k=2 attributes and a convenient single parameter
representation ($\theta$). However for larger k, sets of different
optimal values would be harder to argue about. Also given a set of
real graph snapshots the determination (or approximation) of the
corresponding Markov processes - assuming all properties assumed hold
- is not commented on.

*** 
    The cases where $ k \geq 3$ was not considered in the experiments,
    as we deemed them to be more complex but not more
    difficult. Theoretically, the optimal $\lambda$ for $k \geq 3$ can
    be found for all of the graph invariants by numerical techniques,
    i.e., find the $\lambda$ that maximizes non-central parameters
    $\mu_{\lambda}$ in the cases for size and number of triangles, or
    the ratio $\rho_{\lambda}/\zeta_{\lambda}$ in the case of max
    degree or scan. As for empirical evaluation, one can parametrize 
    the set $\{\lambda \colo \| \lambda \| = 1\}$ for $k \geq 3$
    using hypershperical coordinates and the problem of finding the
    optimal $\lambda$ can be done by exhaustive search. This is 
    inefficient or even intractable for $k \geq 5$, but this is a
    common issue for most global optimization problems.

    The issue of determining or estimating the parameters of the
    underlying Markov chain given a set of snapshots of the graphs is
    an important issue that we failed to comment on explicitly in the
    previous submission. We have added in a paragraph to the revision
    that allude to this issue. In general, it is intractable to
    estimate the parameters of the underlying Markov chains given the
    snapshots. It is possible, however, to estimate the latent vectors
    associated with the first and second-order approximation via
    a spectral embedding technique [1]. We also
    note that more refined models, where the messaging events between each pairs of
    vertices is modeled as point process,
    e.g. [2,3] can also be formulated. The
    snapshot of the graphs is then a binning of these point processes and
    the issue of parameters estimation is fundamental but feasible in 
    these point processes approaches.

[1] Sussman et. al. ``A consistent dot product embedding for
stochastic blockmodel graphs'', Online at http://arxiv.org/abs/1108.2228

[2] Heard et. al. ``Bayesian anomaly detection methods for social
networks'', Annals of Applied Statistics, (4):645--662, 2010.

[3] P. Perry and P. J. Wolfe. ``Point process modeling for directed
interaction networks'', Online at http://arxiv.org/abs/1011.1703v2

***

2. Journal relevance: This is very important since there does not seem
to exist some clear identification of a concrete application from
signal processing. The content would be much more relevant to a
statistics journal or to network analysis audience.

*** 
We have added a new section to motivates the consideration of
time-series of attributed graphs. We believe that the notion of a
time-series is a fundamental concept in signal processing and
furthermore the representation of data as (attributed) graphs is
ubiquitous in many applications domain. Also, anomaly/change-point
detection is one of the main problems in time-series analysis.  As
such, anomaly detection on time-series of graphs is a natural and
challenging problem. There have been recent work on signal processing
on graphs, with many of them on the construction of wavelets on
graphs. Our paper is similar in spirits to these wavelets paper but we
have taken a more statistical point of view, as evidenced by our 
setup of the anomaly detection problem via hypothesis testings along
with the derivations of the limiting behaviour of the test statistics.

Additional Questions:
1. Is the topic appropriate for publication in these transactions?: No

2. Is the topic important to colleagues working in the field?: No (explain):

Explain: As inferred by the comprehensive description of the paper
content included in the comments to authors, there is no explicit link
to the signal processing field e.g. in the form of its corresponding
application potential.

1. Is the paper technically sound?: Yes

why not?:

2. Is the coverage of the topic sufficiently comprehensive and balanced?: Yes

3. How would you describe technical depth of paper?: Appropriate for
the Generally Knowledgeable Individual Working in the Field or a
Related Field

4. How would you rate the technical novelty of the paper?: Not Novel

1. How would you rate the overall organization of the paper?: Satisfactory

2. Are the title and abstract satisfactory?: Yes

Explain:

3. Is the length of the paper appropriate? If not, recommend how the
length of the paper should be amended, including a possible target
length for the final manuscript.: Yes

4. Are symbols, terms, and concepts adequately defined?: Yes

5. How do you rate the English usage? : Satisfactory

6. Rate the Bibliography: Satisfactory

null:

1. How would you rate the technical contents of the paper?: Good

2. How would you rate the novelty of the paper?: Not Novel

3. How would you rate the "literary" presentation of the paper?: Mostly Accessible

4. How would you rate the appropriateness of this paper for
publication in this IEEE Transactions?: Poor Match

Reviewer: 2

Recommendation: RQ - Review Again After Major Changes

Comments:
The paper presents an attribute fusion approach for detecting change
points in dynamic graphs that evolve over time according to a latent
process model. The graphs considered are attributed graphs where each
edge can take one of K possible categories. Four types of graph
invariants for detection are considered, namely the number of edges,
the number of triangles, the maximum degree, and the scan statistic.
Attribute fusion refers to taking non-negative linear combinations of
these invariants computed for each category of edge.

The main contribution of the paper is an asymptotic analysis of the
power of the tests for each invariant as a function of the fusion
parameter lambda, which corresponds to a weighting of the different
attributes. For certain invariants, a closed-form solution for the
optimal lambda is also given. The analysis is performed under a
first-order approximation to the latent process model, which amounts
to a time series of independent Erdos-Renyi graphs (prior to the
change point) and independent graphs sampled from a special instance
of the mixed-membership stochastic blockmodel (after the change
point). Simulation results suggest that the asymptotic estimates of
the power under the first-order approximation are accurate, and that
the optimal choice of fusion parameter lambda under the first-order
approximation is close to the optimal fusion parameter under the exact
model.

The contributions of this paper are timely, relevant, and novel. The
development of signal processing methods for graphs is an important
topic to the signal processing community, as noted by the special
sessions at recent ICASSP and SSP conferences. The literature on
analysis methods for dynamic graphs is limited, and even more so for
attributed graphs. Furthermore, theoretical results that are also
practically useful are extremely rare.

The main weakness of this paper is that it is not organized well and
is not self-contained.

As such, it is difficult to understand the latent process model
without referring to the previous work (reference [1]). There really
is no introduction to the paper, no context is provided for the
problem--aside from the first sentence of the abstract, the reader is
left with no idea why change point detection in dynamic graphs is
useful, and specifically why one may want to use attributed graphs
rather than graphs without attributes. Finally, the latent process
model is a new model, and it would be beneficial to specify early in
the introduction the relationship between this model and other models
that the reader may be familiar with, namely the mixed-membership
stochastic blockmodel and the latent space models of Hoff et al. To
address these weakness, I would suggest to add a few paragraphs
preceding the current section I.A providing context and relationship
to existing models. The authors should find it useful to refer to
other papers published in the Trans on SP to get an idea on the type
of format, organization and style of writing that the TSP reader
expects.

Furthermore, there are terms and quantities that are introduced
without definition, e.g., the bracket notation "[]", e.g., "[K+1]",
the "cadlag process", and the "rdpm" in Sec 1.A. These are not
notations and quantities that are familiar to a signal processing
audience and should be carefully defined before they are introduced.

Specific comments:
- Page 3, figure 1: Please provide a more descriptive caption with
meanings of the regions and trajectories.
- Page 11, line 38: From the definitions of E and F, they appear to be
scalars, but E+F is defined as the convolution of E and F, which does
not make sense.
- Page 13, lines 10-11: What is the change point t* in this example?
- Page 14, line 49: Change "yields" to "yield"
- Page 15, line 4: A plot of each of the normalized test statistics
T_lambda^l (with the optimal lambda for each statistic) against time
should also be included so the reader can get a sense for how
anomalous this particular time step happens to be.
- Page 17, figure 6: Why are there multiple lines for the second order
approximation and exact model? Are they for different values of r? If
so, the values of r should be indicated somewhere in the plots
themselves (perhaps by reducing the number of values of r shown) or by
listing them in the caption.
- Appendix, page 1, line 31: Change "U-statistics" to "U-statistic"

Additional Questions:
1. Is the topic appropriate for publication in these transactions?: Yes

2. Is the topic important to colleagues working in the field?: Yes

Explain:

1. Is the paper technically sound?: Yes

why not?:

2. Is the coverage of the topic sufficiently comprehensive and
balanced?: Important Information is missing or superficially treated.

3. How would you describe technical depth of paper?: Suitable only for an expert

4. How would you rate the technical novelty of the paper?: Somewhat Novel

1. How would you rate the overall organization of the paper?: Poor

2. Are the title and abstract satisfactory?: Yes

Explain:

3. Is the length of the paper appropriate? If not, recommend how the
length of the paper should be amended, including a possible target
length for the final manuscript.: Yes

4. Are symbols, terms, and concepts adequately defined?: No

5. How do you rate the English usage? : Satisfactory

6. Rate the Bibliography: Satisfactory

null:

1. How would you rate the technical contents of the paper?: Excellent

2. How would you rate the novelty of the paper?: Sufficiently Novel

3. How would you rate the "literary" presentation of the paper?:
Partially Accessible

4. How would you rate the appropriateness of this paper for
publication in this IEEE Transactions?: Weak Match