### 2013-02-27 decision2.tex
Version accepted to TSP

Thank you for the careful reading of our manuscript and the referee
report and we deeply appreciate your comments and critiques on the
manuscript. The specific comments and suggestions for review have been
followed, and the result is an improved manuscript. In particular,
the discussions surrounding the theoretical results in Section IV and
the experiments in Section V had been expanded, most importantly the
addition of Figure 5, illustrating, through Q-Q plots, 
the convergence of the statistics to their limiting distribution. Below we
give specific responses to the comments of the reviewers.  Again, we
thank the editor and reviewers for their valuable comments.

==================================================
Reviewer: 1

Recommendation: AQ - Publish In Minor, Required Changes

Comments:
(1) Expand the discussion on the connection of your approach with the
wavelet representations for graphs; a small paragraph, 3-4 sentences,
would suffice.

*** Response: 
We have expanded the discussion on related works. In
particular, we described the similarities between the construction and
use of the wavelet representation for traffic analysis in [Crovella & Kolaczyk, 2003] and
anomaly detection using locality/scan statistics in our manuscript.

(2) Comment more explicitly on the choice of the running-average window l. Is there a guideline?

*** Response: 
We had added a comment about the choice of $l$ right before
the end of Section 3. In general, the choice of $l$ depends on the
practitioner. Small values of $l$ could
induce large fluctuations in the temporal standardization statistics,
which might lead to a higher false positive rate. Large values of $l$,
on the other hand, could induce over-smoothing, leading to a higher
false negative rate. As for the Enron data
experiment, $l = 20$ was used in the earlier scan paper on Enron graph
(Priebe et al. 2005) so we used it to keep the comparisons between the
two papers meaningful.

(3) Add reference [*] [*] C. E. Priebe, N. H. Lee, Y. Park, and
M. Tang, "Attribute Fusion in a Latent Process Model for Time Series
of Graphs," 2011 IEEE Workshop on Statistical Signal Processing
(SSP2011), July 2011.
*** Response: This had been done.

==================================
Reviewer: 2

Recommendation: RQ - Review Again After Major Changes

Comments:

(1) The paper has been improved somewhat but it is still written in a
terse mathematical style that is not consistent with the signal
processing audience. The introductory section is from [1] but the
reader would not know this if [1] had not been read.  The Lemmas and
Theorems in the power analysis section are inadequately discussed in
terms of their practical applicability to problems of signal
processing and time series analysis. In particular, the weak
convergence theory of Thms 4, 6 and 9 is interesting but the utility
of these theorems is not adequately illustrated.  The Gumbel limits
are not explained, interpreted or illustrated in the numerical
section. The authors could, for example, illustrate how these results
can be used to set thresholds or select the linear parameter
lambda. How large does n and l have to be for these limits to become
good apprxoimation? Q-Q plots of simulation or real (Enron) data
against student-t and Gumbel would give an idea of how fast these
approximations converge.  In any case, any theoretical results that
are not illustrated for a SP problem should be removed from the paper.

*** Response 
We have reorganized and expanded the paper to make the connections
between the theoretical results in Section IV and their applicability
to signal processing and time-series analysis more explicit. In
particular, section IV now starts off with a discussion relating the
change-point detection framework in Section III with the theoretical
results, i.e., the theoretical results provided power estimates for
each choice of the fusion parameter $\lambda$. We have also note how
the theoretical results in this paper are extensions of results that 
exist in the literature.

The rate of convergence for the limiting distribution is an important
result that we had ignored in the paper. The paper of [1] contains
evaluation (via Monte Carlo estimates) of rate of convergence of size,
max degree, number of triangles and scan to the corresponding limiting
distributions, i.e., normal, Gumbel, normal and Gumbel in the
Erdos-Renyi and kindey-egg settings. Note that they consider only a
single (unattributed) graph at a time. The paper Pao, Coppersmith and
Priebe (citation below) showed that the limiting distribution with the
associated parameters is sufficiently accurate for the size and max
degree statistic starting at $n = 100$ and for number of of triangles
and scan statistics starting at $n = 1000$.

In general, for the case of degree and number of triangles provided
that the normal distribution is a good large-sample approximation,
then the weak convergence of the temporal standardization statistics
to the Student's t-distribution holds independent of $l$, as long as
the underlying time-series of graphs is stationary during that time
interval.

For the case of max degree and scan, the weak convergence of
their temporal standardization to the Gumbel distribution requires
both that $n$ is sufficiently large (so that the Gumbel distribution
is a good large-sample approximation at any time $t$) and that $l$ is
sufficiently large, e.g., $l \geq 100$. However, if $n$ is
sufficiently large, then the limiting distribution of the temporal
standardization for any value of $l \geq 2$ converges weakly to the
convolution of $l$ Gumbel distributed random variables as long as the
underlying time-series is stationary during that time interval. Power
estimates in this case (where both the null and alternative is a
convolution of Gumbel random variables) is also straightforward but tedious.

The manuscript has now been revised to include parts of the above
discussion on the limiting distribution (right after the statement of
the relevant results) and how it is applied to the simulation
experiment setup (second paragraph of Section V). In particular, Q-Q
plots for the temporal standardization of the max degree, number of
triangles and scan statistics against the limiting distributions had
been added to the manuscript. These Q-Q plots help to illustrate the
convergence of these temporal standardization statistics to their
limiting distribution.

[1] Pao, Coppersmith, and Priebe, ``Statistical inference of random graphs: Comparative
power analysis via Monte Carlo'' Journal of Computational and Graphical Statistics, 2011.

***

(2) The Introduction should very clearly state the differences between [1]
and the present work. The reader should not have to read [1], which
this reviewer did, in order to figure this out.

*** Response
The manuscript now include a short paragraph at the end of the introductory
section describing the relationships between [1], another paper that
was an earlier version of the manuscript and was presented at the 2011
SSP workshop in Nice, France, and the manuscript. In particular the
current manuscript used the generative model for the time-series of graphs
from [1]. Also, the setup of anomaly/change-point detection and the
use of temporal standardization of the graph invariants as the test
statistics are also identical to that done in [1] (this setup or
its variation had been used previously by some of the authors, see
e.g., Priebe et al ``Scan statistics on Enron graphs''). However, [1] only
considered the use of the size invariant. The earlier version of the
manuscript that was presented at the 2011 SSP workshop motivates the
current manuscript and include a preliminary version of the simulation
experiment in Section V.  
***

(3) In Fig 4 the 1st and 2nd order approximations appear to upper bound
the exact curve for all cases studied.  Is this provable? The authors
should comment.

*** Response
Two of the authors of the manuscript had noticed the above
phenomenon (and the fact that the 2nd-order approximation is monotone
increasing in $r$, the process rate parameter) in their previous paper
[1]. However, they have not been able to explain the phenomenon. It
might be possible to show that the
second-order approximations is monotone increasing in $r$, which also implies
that the first-order approximation is an upper bound for the second order
approximations, for the size and number of triangle statistic using some
recent results in [2]. However, to the best of our knowledge,
the explanation for the full phenomenon is still elusive. We would be
very much interested in knowing an explanation behind this phenomenon.

[2] Priebe and Rukhin, ``Invariant theory for hypothesis testing on
graphs'', Proceedings of the 58th World Statistic Congress, 2011. 
***

(4) Non-standard notation (e.g. [n], [K+1]) has now been defined but
it should be aligned with a more standard notation (e.g. {1,...,n} and
{1,...,K+1}) familar to SP audience. There are also many typographical
errors in the manuscript. The authors need to go through and carefully
correct errors of syntax and omissions of important assumptions. A
sampling of the ones I found is:

* Eqn (2) , <x_u,x_v> should be <X_u,X_v>.
*Line 14 on p. 5: "phenoma" should be "phenomenon"
*Line 19 on p. 5; "Permutation of the vertex labels does not affect
our subsequent analysis"
*Line 23 p 5: "Of general interests" should be replaced by "Of general interest"
*Line 31 on p 7: "parameters" should be "parameter"
*Line 4 on p 8: "interests" should be "interest"
*Line 12 on p 8: the "I" notation used to define \Gamma_{uv}(k) shoudl
be defined explicitly as a vector of indicator functions since this is
not standard notation. Furthermore, on the next line the statement
\Gamma_{uv}=0 is confusing since \Gamma_{uv} is a vector. A better
notation will be \Gamma_{uv} =[0,\ldots,0]^T.
*Thm 1 needs to be stated more precisely with all necessary
assumptions, as in [1]. The quantites defined are functions of m,n,l
and involve matrices eta_{ij}. In [1] these matrices are assumed to be
positive definite and the matrix \zeta is assumed to be
non-zero. These assumptions need to be stated here too.

*** Response
The above errors had all been fixed. In particular the statement of
the theoretical results had been revised or corrected. An additional
assumption regarding the matrix \zeta being positive definite is now
included in Theorem 1. All occurrences of the fraction $\sqrt{\pi}/6$ in Theorem 6 and
Theorem 9 had now been changed/corrected to $\sqrt{6}/pi$.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: