### 2013-02-27 response2.tex

\documentclass{article}
\pdfminorversion=4
\usepackage{parskip}
\usepackage{fontenc}
\begin{document}
\title{Response to reviewers}
\maketitle
\date{}
Thank you for the careful reading of our manuscript and the referee
report and we deeply appreciate your comments and critiques on the
manuscript. The specific comments and suggestions for review have been
followed, and the result is an improved manuscript. In particular,
the discussions surrounding the theoretical results in Section IV and
the experiments in Section V had been expanded, most importantly the
addition of Figure 5, illustrating, through Q-Q plots, 
the convergence of the statistics to their limiting distribution. Below we
give specific responses to the comments of the reviewers (the
reviewers comments are in italic). Again, we
thank the editor and reviewers for their valuable comments.

\section{Comments from Reviewer 1}
Recommendation: AQ - Publish In Minor, Required Changes

Comments:
{\em (1) Expand the discussion on the connection of your approach with the
wavelet representations for graphs; a small paragraph, 3-4 sentences,
would suffice.}

*** Response: We have expanded the discussion on related works. In
particular, we described the similarities between the construction and
use of the wavelet representation for traffic analysis in [Crovella \&
Kolaczyk, 2003] and anomaly detection using locality/scan statistics
in our manuscript.

{\em (2) Comment more explicitly on the choice of the running-average
window l. Is there a guideline?}

*** Response: We had added a comment about the choice of $l$ right
before the end of Section 3. In general, the choice of $l$ depends on
the practitioner. Small values of $l$ could induce large fluctuations
in the temporal standardization statistics, which might lead to a
higher false positive rate. Large values of $l$, on the other hand,
could induce over-smoothing, leading to a higher false negative
rate. As for the Enron data experiment, $l = 20$ was used in the
earlier scan paper on Enron graph (Priebe et al. 2005) so we used it
to keep the comparisons between the two papers meaningful.

{\em (3) Add reference [*] C. E. Priebe, N. H. Lee, Y. Park, and
M. Tang, "Attribute Fusion in a Latent Process Model for Time Series
of Graphs," 2011 IEEE Workshop on Statistical Signal Processing
(SSP2011), July 2011.}

*** Response: This had been done.

\section{Comments from Reviewer 2}

Recommendation: RQ - Review Again After Major Changes

Comments:

{\em (1) The paper has been improved somewhat but it is still written in a
terse mathematical style that is not consistent with the signal
processing audience. The introductory section is from [1] but the
reader would not know this if [1] had not been read.  The Lemmas and
Theorems in the power analysis section are inadequately discussed in
terms of their practical applicability to problems of signal
processing and time series analysis. In particular, the weak
convergence theory of Thms 4, 6 and 9 is interesting but the utility
of these theorems is not adequately illustrated.  The Gumbel limits
are not explained, interpreted or illustrated in the numerical
section. The authors could, for example, illustrate how these results
can be used to set thresholds or select the linear parameter
lambda. How large does n and l have to be for these limits to become
good apprxoimation? Q-Q plots of simulation or real (Enron) data
against student-t and Gumbel would give an idea of how fast these
approximations converge.  In any case, any theoretical results that
are not illustrated for a SP problem should be removed from the paper.}

*** Response: We have reorganized and expanded the paper to make the
connections between the theoretical results in Section IV and their
applicability to signal processing and time-series analysis more
explicit. In particular, section IV now starts off with a discussion
relating the change-point detection framework in Section III with the
theoretical results, i.e., the theoretical results provided power
estimates for each choice of the fusion parameter $\lambda$. We have
also note how the theoretical results in this paper are extensions of
results that exist in the literature.

The rate of convergence for the limiting distribution is an important
result that we had ignored in the paper. The paper of [1] contains
evaluation (via Monte Carlo estimates) of rate of convergence of size,
max degree, number of triangles and scan to the corresponding limiting
distributions, i.e., normal, Gumbel, normal and Gumbel in the
Erdos-Renyi and kindey-egg settings. Note that they consider only a
single (unattributed) graph at a time. The paper Pao, Coppersmith and
Priebe (citation below) showed that the limiting distribution with the
associated parameters is sufficiently accurate for the size and max
degree statistic starting at $n = 100$ and for number of of triangles
and scan statistics starting at $n = 1000$.

In general, for the case of degree and number of triangles provided
that the normal distribution is a good large-sample approximation,
then the weak convergence of the temporal standardization statistics
to the Student's t-distribution holds independent of $l$, as long as
the underlying time-series of graphs is stationary during that time
interval.

For the case of max degree and scan, the weak convergence of
their temporal standardization to the Gumbel distribution requires
both that $n$ is sufficiently large (so that the Gumbel distribution
is a good large-sample approximation at any time $t$) and that $l$ is
sufficiently large, e.g., $l \geq 100$. However, if $n$ is
sufficiently large, then the limiting distribution of the temporal
standardization for any value of $l \geq 2$ converges weakly to the
convolution of $l$ Gumbel distributed random variables as long as the
underlying time-series is stationary during that time interval. Power
estimates in this case (where both the null and alternative is a
convolution of Gumbel random variables) is also straightforward but tedious.

The manuscript has now been revised to include parts of the above
discussion on the limiting distribution (right after the statement of
the relevant results) and how it is applied to the simulation
experiment setup (second paragraph of Section V). In particular, Q-Q
plots for the temporal standardization of the max degree, number of
triangles and scan statistics against the limiting distributions had
been added to the manuscript. These Q-Q plots help to illustrate the
convergence of these temporal standardization statistics to their
limiting distribution.

[1] Pao, Coppersmith, and Priebe, ``Statistical inference of random
graphs: Comparative power analysis via Monte Carlo'' Journal of
Computational and Graphical Statistics, 2011.

{\em (2) The Introduction should very clearly state the differences between [1]
and the present work. The reader should not have to read [1], which
this reviewer did, in order to figure this out.}

*** Response: The manuscript now include a short paragraph at the end
of the introductory section describing the relationships between [1],
another paper that was an earlier version of the manuscript and was
presented at the 2011 SSP workshop in Nice, France, and the
manuscript. In particular the current manuscript used the generative
model for the time-series of graphs from [1]. Also, the setup of
anomaly/change-point detection and the use of temporal standardization
of the graph invariants as the test statistics are also identical to
that done in [1] (this setup or its variation had been used previously
by some of the authors, see e.g., Priebe et al ``Scan statistics on
Enron graphs''). However, [1] only considered the use of the size
invariant. The earlier version of the manuscript that was presented at
the 2011 SSP workshop motivates the current manuscript and include a
preliminary version of the simulation experiment in Section V.  ***

{\em (3) In Fig 4 the 1st and 2nd order approximations appear to upper bound
the exact curve for all cases studied.  Is this provable? The authors
should comment.}

*** Response: Two of the authors of the manuscript had noticed the
above phenomenon (and the fact that the 2nd-order approximation is
monotone increasing in $r$, the process rate parameter) in their
previous paper [1]. However, they have not been able to explain the
phenomenon. It might be possible to show that the second-order
approximations is monotone increasing in $r$, which also implies that
the first-order approximation is an upper bound for the second order
approximations, for the size and number of triangle statistic using
some recent results in [2]. However, to the best of our knowledge, the
explanation for the full phenomenon is still elusive. We would be very
much interested in knowing an explanation behind this phenomenon.

[2] Priebe and Rukhin, ``Invariant theory for hypothesis testing on
graphs'', Proceedings of the 58th World Statistic Congress, 2011. 
***

{\em (4) Non-standard notation (e.g. [n], [K+1]) has now been defined but
it should be aligned with a more standard notation (e.g. {1,...,n} and
{1,...,K+1}) familar to SP audience. There are also many typographical
errors in the manuscript. The authors need to go through and carefully
correct errors of syntax and omissions of important assumptions. A
sampling of the ones I found is:

\begin{enumerate}
  \item Eqn (2) , $<x_u,x_v>$ should be $<X_u,X_v>$.
  \item Line 14 on p. 5: "phenoma" should be "phenomenon"
  \item Line 19 on p. 5; "Permutation of the vertex labels does not affect
our subsequent analysis"
\item Line 23 p 5: "Of general interests" should be replaced by "Of general interest"
\item Line 31 on p 7: "parameters" should be "parameter"
\item Line 4 on p 8: "interests" should be "interest"
\item Line 12 on p 8: the "I" notation used to define $\Gamma_{uv}(k)$ should
be defined explicitly as a vector of indicator functions since this is
not standard notation. Furthermore, on the next line the statement
$\Gamma_{uv}=0$ is confusing since $\Gamma_{uv}$ is a vector. A better
notation will be $\Gamma_{uv} =[0,\ldots,0]^T$.
\item Thm 1 needs to be stated more precisely with all necessary
assumptions, as in [1]. The quantites defined are functions of m,n,l
and involve matrices $eta_{ij}$. In [1] these matrices are assumed to be
positive definite and the matrix $\zeta$ is assumed to be
non-zero. These assumptions need to be stated here too.
\end{enumerate}}

*** Response: The above errors had all been fixed. In particular the
statement of the theoretical results had been revised or corrected. An
additional assumption regarding the matrix $\zeta$ being positive
definite is now included in Theorem 1. All occurrences of the fraction
$\sqrt{\pi}/6$ in Theorem 6 and Theorem 9 had now been
changed/corrected to $\sqrt{6}/pi$.
\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: