### 2013-02-27 ssp2011.tex

\documentclass[10pt,draftclsnofoot,onecolumn]{IEEEtran}
\usepackage{fontenc}
\pdfminorversion=4
%\usepackage{lmodern}
%\usepackage[utopia]{mathdesign}
\usepackage{graphicx}
%\usepackage{xltxtra}
%\setmainfont[Mapping=tex-text]{Linux Libertine}
\usepackage[cmex10]{amsmath}
\interdisplaylinepenalty=2500
\usepackage{array}
\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage[tight,normalsize,sf,SF]{subfigure}
\ifCLASSOPTIONcompsoc
\usepackage[font=normalsize,labelfont=sf,textfont=sf]{subfig}
\else
\usepackage[font=footnotesize]{subfig}
\fi
\usepackage{subfig}
\usepackage{parskip}
\usepackage{mathrsfs}
\usepackage{bm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
%\usepackage[colorlinks=true,pagebackref,linkcolor=magenta]{hyperref}
\usepackage[noadjust]{cite}
\renewcommand\arraystretch{1.2}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\def\mathllap{\mathpalette\mathllapinternal}
\def\mathllapinternal#1#2{%
\llap{$\mathsurround=0pt#1{#2}$}% $
}
\def\clap#1{\hbox to 0pt{\hss#1\hss}}
\def\mathclap{\mathpalette\mathclapinternal}
\def\mathclapinternal#1#2{%
\clap{$\mathsurround=0pt#1{#2}$}%
}
\def\mathrlap{\mathpalette\mathrlapinternal}
\def\mathrlapinternal#1#2{%
\rlap{$\mathsurround=0pt#1{#2}$}% $
}
\bibliographystyle{IEEEtran}
\begin{document}
\title{Attribute fusion in a latent process model for time series of
  graphs}
\author{Minh~Tang, Youngser~Park, Nam~H.~Lee, and Carey~E.~Priebe*%
  \thanks{Copyright (c) 2012 IEEE. Personal use of this material is permitted. However, permission to use this material for any other purposes must be obtained from the IEEE by sending a request to pubs-permissions@ieee.org.}
\thanks{Minh~Tang is with the Department of Applied Mathematics and
  Statistics, Johns Hopkins University, Baltimore, MD 21218
  USA. (email: mtang10@jhu.edu)}%
\thanks{Youngser~Park is with the Center of Imaging Science, Johns
  Hopkins University, Baltimore, MD 21218
  USA. (email:youngser@jhu.edu)}%
\thanks{Nam~H.~Lee is with the Department of Applied Mathematics and
  Statistics, Johns Hopkins University, Baltimore, MD 21218
  USA. (email: nhlee@jhu.edu)}%
\thanks{Carey~E.~Priebe is with the Department of Applied Mathematics and
  Statistics, Johns Hopkins University, Baltimore, MD 21218
  USA. (phone: 410-516-7200; fax:410-516-7459; email: cep@jhu.edu)}}%
\markboth{IEEE Transactions on Signal Processing}{}
\maketitle
\begin{abstract}
  Hypothesis testing on time series of attributed graphs has
  applications in diverse areas, e.g., social network analysis
  (wherein vertices represent individual actors or organizations),
  connectome inference (wherein vertices are neurons or brain regions)
  and text processing (wherein vertices represent authors or
  documents).  We consider the problem of anomaly/change point
  detection given the latent process model for time series of graphs
  with categorical attributes on the edges presented in \cite{lee11}.
  Various attributed graph invariants are considered, and their power
  for detection as a function of a linear fusion parameter is
  presented.  Our main result is that inferential performance in
  mathematically tractable first-order and second-order approximation
  models does provide guidance for methodological choices applicable
  to the exact (realistic but intractable) model. Furthermore, to the
  extent that the exact model is realistic, we may tentatively
  conclude that approximation model investigations have some bearing
  on real data applications.
\end{abstract}
 \begin{IEEEkeywords}
   Anomaly detection, Attributed Random Graphs, Random Dot
   Product Graphs.
 \end{IEEEkeywords}
%\begin{center} \bfseries EDICS Category: SSP-SSAN, SSP-DETC \end{center}
%\pagebreak
\section{Introduction}
\label{sec:introduction}
The notion of data as a time series is fundamental to signal
processing. However, time-series analysis had been mostly confined to
data in Euclidean spaces, while the representation of data as graphs
with the vertices as the entities and the edges as the relationships
between the entities is now ubiquitous in many application domains,
e.g., social networks (wherein vertices represent individual actors or
organizations), neuroscience (wherein vertices are neurons or brain
regions), and text processing (wherein vertices represent authors or
documents). Furthermore, in many of these application domains, the
edges between the vertices, in addition to representing the presence
of a relationship, can also be used to identify the attributes
associated with that relationship. For example, in social network
graphs where edges represent email communications, the attributes
signify the context of the email messages. Finally, in many of these
application domains, the underlying data changes over time, and thus a
representation of data as a time-series of attributed graphs is
natural. The analysis of time-series of (attributed) graphs thus
presents a natural and challenging extension to the signal processing
community.

One of the main exploitation tasks in time-series analysis is the
problem of anomaly/change-point detection, where anomaly is broadly
interpreted to mean deviation from some ``normal'' pattern and
change-point is the time window at which the anomalous deviation
occurs. In the context of a time-series of graphs, an anomaly can be
for example the presence of a specific subgraph, or frequent occurrences
of some subgraphs, or other more vague notions such as existence of a
community structure.

The notion of signal processing for graphs appears to be in an
emerging state, with significant research activity in constructing
wavelet representations for graphs, e.g.,
\cite{hammond10:_wavel,r.coifman06:_diffus,crovella03:_graph,narangss:_perfec}.
The formulation of graph analysis problems in a signal processing
framework is also of potential interest, e.g,
\cite{miller10:_towar_euclid}, wherein a subgraph detection problem is
presented, test statistics are defined, noise and signal models are
introduced, and ROC analysis is performed. The construction of
wavelets representations on graphs is analogous to the classical
wavelets constructions, but adapted to use the topology of the graphs
instead of the usual Euclidean space topology. For example, the
wavelet representations of \cite{hammond10:_wavel} and
\cite{r.coifman06:_diffus} are constructed using the combinatorial and
normalized Laplacian matrices of the graphs, respectively, while
\cite{crovella03:_graph} defined wavelets via the $n$-hop
neighborhoods. The wavelet construction in \cite{crovella03:_graph}
and its use in {\em spatial} traffic analysis (emphasis in 
\cite{crovella03:_graph}) is similar to the use of locality/scan
statistics for anomaly detection in
\cite{priebe05:_scan_statis_enron_graph} and the current
paper. Indeed, the locality statistic at a vertex $v$ is the size of an
$n$-hop neighborhood starting at $v$ defined via shortest path
distance and the scan statistic is the maximum of the locality
statistics over all the $n$-hop neighborhoods. Traffic analysis by
tracking the coefficients of the wavelet representations is similar to
anomaly detection by tracking the values of the scan statistics.

In this paper, we investigate the anomaly/change-point detection
problem for time-series of attributed graphs. In the spirit of
time-series analysis, we approach this problem via test statistics
based on the moving average of some graph invariants. We derive
limiting distributions for these test statistics under a generative
model for the attributed graphs. The experimental evidences indicate that these
test statistics have sufficient power against the null hypothesis for
a general class of change-point problem but that many interesting
questions remain to ponder.

The current paper builds upon ideas previously presented in
\cite{lee11} and \cite{priebe11:_attrib}, among others. The generative
model for time-series of attributed graphs used in this paper is from
\cite{lee11}. Hypothesis testing for time-series of graphs and
attribute fusion using a linear fusion parameter $\lambda$ were
considered in \cite{lee11,priebe11:_attrib}; limiting distribution for
the size graph invariant was given in \cite{lee11}. A version of the
experimental results for simulated data included in this paper was first
presented in \cite{priebe11:_attrib}.

The paper is structured as follows. We present an overview of the
generative model for time-series of attributed graphs \cite{lee11} in
\S~\ref{sec:latent-process-model}. A change-point detection problem
along with some relevant test statistics are described in
\S~\ref{sec:change-point-detect}. The distributions of these test
statistics and the power estimates for hypothesis testing using
these test statistics are given in \S~\ref{sec:power-estimates}. 
\S~\ref{sec:inference-examples} presents a
synthetic data example and an example using the Enron email corpus to
illustrate our methodology.

\section{Latent Process Model}
\label{sec:latent-process-model}
The abundance of graph representations in diverse application domains
leads to a proliferation of random graph models to model the
association among the entities. In addition to the classical
Erd\"{o}s-R\'{e}nyi model \cite{Erdos1959}, there is the exponential
graph models \cite{holland81}, the stochastic block model
\cite{Holland1983,Wang1987}, the latent position model
\cite{hoff02:_laten}, the dot product model for unattributed graphs
\cite{young07:_random} and the latent process model for attributed
graphs \cite{lee11}. We now describe briefly the latent position model
of \cite{hoff02:_laten} and the dot product model of
\cite{young07:_random} as they are closely related to the latent
process model in \cite{lee11}. The latent process model serves as
the generative model for our attributed graphs and thus plays a
key role in our setup of the anomaly detection problem in 
\S~\ref{sec:change-point-detect}.

The latent space model is one where each vertex is associated with a
latent random vector. There may also be additional covariates
information but those are not relevant to the current exposition. The
vectors are independent and identically distributed and the
probability of an edge between two nodes depends only on their latent
vectors. Conditioned on the latent vectors, the presence of each edge
is an independent Bernoulli trial. For example, if $p_{uv}$ is the
probability of an edge between the vertices $u$ and $v$, then $p_{uv}
= \tfrac{1}{1 + \exp( - \|z_u - z_v\|)}$ where $z_u$ and $z_v$ are the
latent vectors associated with vertex $u$ and $v$, respectively.
Another example of a latent space model is the random dot product
graph model \cite{young07:_random}. Under the random dot product graph
model, the probability of an edge between two vertices is given
by the dot product of their respective latent vectors, i.e., $p_{uv} =
\langle z_u, z_v \rangle$. A motivation behind the random dot
product graph model is that, in a social network with edges
indicating friendships, the components of the vector may be
interpreted as the relative interest of the individual in various
topics. The magnitude of the vector can be interpreted as how
talkative the individual is, with more talkative individuals more
likely to form relationships. Talkative individuals interested in the
same topics are most likely to form relationships while individuals
who do not share interest are unlikely to form relationships.

The latent process model for time-series of attributed graphs was
presented in \cite{lee11}. The main ideas underlying the model is as
follows. The model is motivated by the assumption that each vertex is
associated with a stochastic process, in this case a finite-state
continuous time Markov chain. The stochastic processes then associate
to each vertex $v$, for the time interval $(t-1,t]$, a (latent) random
vector $X_v(t)$ in the unit simplex with the $k$-th entry of $X_v(t)$
denoting the proportion of time the stochastic process spent in state
$k$. The probability of interactions between vertices $u$
and $v$ during $(t-1, t]$ is then given by the dot product of $X_u$
and $X_v$ a la the dot product model of \cite{young07:_random}, with a
slight modification to account for the presence of attributes on the
edges.

We now give the necessary details to make the above ideas precise. For
any positive integer $n$, we shall denote by $[n]$ the set of numbers
$\{1,2,\dots,n\}$. Let $G = (V,E)$ be an undirected graph, equipped
with an edge-attribution function $\phi \colon \tbinom{V}{2} \mapsto
[K+1]$ for some positive integer $K$. We shall require that $\phi$
satisfies $\phi(e) \leq K$ for $e \in E$ and $\phi(e) \equiv K+1$ for
$e \not \in E$. We refer to $\phi(e)$ as the attribute of edge $e \in
E$, and the pair $(G,\phi)$ as an attributed graph. Let $\mathscr{S}$
be the unit simplex in $\mathbb{R}^{K}$, i.e.,
\begin{equation}
  \mathscr{S} = \{ \xi \in [0,1]^{K}
  \colon \sum_{k = 1}^{K} \xi_k \leq 1 \}.
\end{equation}

A {\em random dot product space} for attributed graphs with vertices
in $[n]$ and edge attributes $[K]$ is a pair $(\mathbf{X},G)$ of
random elements such that
\begin{enumerate}
\item $\mathbf{X} = \{X_v\}_{v = 1}^{n}$ is a collection of
  independent $\mathscr{S}$-valued random vectors.
\item $G$ is a random undirected graph with vertex set $[n]$ such
  that the edges of $G$ are conditionally independent given
  $\mathbf{X}$ and that
  \begin{equation}
    \label{eq:1}
    \mathbb{P}(u \sim v \,|\, \mathbf{X}) = \mathbb{P}(u \sim v \, |
    \, X_u,X_v) = \langle X_u, X_v \rangle.
  \end{equation}
 Furthermore, we also have 
  \begin{equation}
    \label{eq:11}
    \mathbb{P}(\phi(\{u,v\}) = k \, | \, u \sim
    v) = \frac{X_{u,k} X_{v,k}}{\langle X_u, X_v \rangle}
  \end{equation}
  i.e., the pair $\{u,v\}$ is an edge in $G$ having attribute $k$ with
  probability $X_{u,k} X_{v,k}$.  
\end{enumerate}
We say that a c\'{a}dl\'{a}g (right continuous with left limit)
process $W \colon [0,\infty) \mapsto [K+1]^{n}$ induces the sequence
of random dot product spaces $\mathscr{V} = (\mathbf{X}(t), G(t))_{t =
1}^{\infty}$ if
\begin{enumerate}
\item Each $(\mathbf{X}(t), G(t))$ is a random dot product space with
  vertices $[n]$ and edge attributes $[K]$. Furthermore, for each $v
  \in [n]$, $k \in [K]$ and $t \in \mathbb{N}$, we have
  \begin{equation}
    \label{eq:54}
  X_{v,k}(t)  = \int_{t - 1}^{t}{ \mathbf{1}\{W_v(\omega) = k\}\, d\omega}
  \end{equation}
  where we denote $W_{v}(\omega)$ as the $v$-th component of $W(\omega)$ and
  $X_{v,k}$ as the $k$-th component of $X_v$.  
\item  For each $t \in \mathbb{N}$, we have
  \begin{equation}
    \label{eq:2}
    \mathbb{P}(G(t) = g \,|\, \mathscr{F}_{\leq t}) = \mathbb{P}(G(t)
    = g \,|\, \mathbf{X}(t))
  \end{equation}
where $\mathscr{F}_{\leq t}$ is the $\sigma$-field generated by $\{W(s)
  \colon s \leq t\}$.
\end{enumerate}
We will call any pair $(\mathscr{V}, W)$ that satisfies the above
properties a random dot process model.

The random dot process model can be interpreted as follows. During
each time interval $(t-1,t]$, a particular vertex is inclined to
communicate with other vertices on a variety of topics (the topics are
enumerated as the elements of $[K]$). As the inclination of a vertex
$v$ is governed by its underlying stochastic process $W_v$, we can
view the $\{W_v\}$ as encapsulating a notion of changes in the
inclination of the vertices over time. The propensity for
vertex $v$ to communicate about a particular topic $k$ is $X_{v,k}(t)$
and two vertices $u$ and $v$ are more likely to exchange communication
on topic $k$ if the propensity to communicate about topic $k$ is high
for both $u$ and $v$. Eq.~\eqref{eq:1} and Eq.~\eqref{eq:11} attempt
to capture these two assumptions.

In this paper, we are interested in the pairs $(\mathscr{V}, W)$
possessing the following properties:
\begin{enumerate}
\item For each $t \in \mathbb{N}$ and vertex $v \in [n]$, there exists
  a matrix $\mathbf{Q}^{(v)}(t)$ such that $W_v$, when restricted to
  the interval $[t, t+1)$, is a stationary, continuous-time Markov
  chain with state space $[K+1]$, intensity matrix
  ${\mathbf{Q}^{(v)}(t)}$, and stationary distribution
  $\pi^{(v)}(t)$. $W$ is thus a piece-wise stationary, continuous-time
  Markov chain with state space $[K+1]^{n}$ and intensity matrix
  $\otimes_{v=1}^{n}\mathbf{Q}^{(v)}(t)$.
\item There exists a $t^{*} \in \mathbb{N}$ and a $m < n$ such that,
  for some $\pi_0, \pi_1, \mathbf{Q}_0$ and $\mathbf{Q}_1$ we have
  \begin{enumerate}
  \item for $t < t^{*} - 1$,
    \begin{gather*}
      \pi^{(1)}(t) = \dots = \pi^{(n)}(t) = \pi_0 \\
      \mathbf{Q}^{(1)}(t) = \dots = \mathbf{Q}^{(n)}(t) = \mathbf{Q}_0,
    \end{gather*}
  \item  for $t \geq t^{*} - 1$,
    \begin{gather*}
      \pi^{(1)}((t)) = \dots = \pi^{(m)}(t) = \pi_1 \\
      \pi^{(m+1)}(t)  = \dots = \pi^{(n)}(t) = \pi_0 \\
      \mathbf{Q}^{(1)}(t)  = \dots = \mathbf{Q}^{(m)}(t) = \mathbf{Q}_1 \\
      \mathbf{Q}^{(m+1)}(t) = \dots = \mathbf{Q}^{(n)}(t) = \mathbf{Q}_0 
    \end{gather*}
  \end{enumerate}
\end{enumerate}
The above properties characterize a random dot process model with a
change-point phenomenon for $(X(t),G(t))$ at time $t = t^{*}$. We chose
$\{1,2,\dots,m\}$ as the set of vertices with a change in the
stationary distribution and intensity matrix at time $t = t^{*} - 1$
for ease of discussion. Permutations of the vertex labels do not
affect our subsequent analysis. We will refer to $(t^{*}, m, \pi_0,
\pi_1, \mathbf{Q}_0, \mathbf{Q}_1)$ as the change-parameters.  Because
$W$ is completely determined by the change-parameters, we often choose
to omit $W$ and only mention $\mathscr{V}$ when referring to random
dot process models with change-point phenomenon. For a random dot
process model
$\mathscr{V}$ with change-parameters $(t^{*}, m, \pi_0, \pi_1,
\mathbf{Q}_0, \mathbf{Q}_1)$, we can construct several approximations
to $\mathscr{V}$. Of particular interest are the following two
approximations.
\begin{definition}
  \label{def:1}
  Let $\mathscr{V}$ be a random dot process model (rdpm) with change
  parameters $(t^{*}, m, \pi_0, \pi_1, \mathbf{Q}_0,
  \mathbf{Q}_1)$. The first order approximation $\bar{\mathscr{V}}$ of
  $\mathscr{V}$ is the sequence $\{(\bar{X}(t), \bar{G}(t)\}_{t =
    1}^{\infty}$ of {\em independent} random dot product spaces such
  that
 \begin{enumerate}
 \item for $t < t^{*} - 1$,
   \begin{equation}
     \label{eq:5}
     \bar{X}_{v}(t)  \equiv \hat{\pi}_0 \quad \text{for all $v \in [n]$}
   \end{equation}
 \item for $t \geq t^{*} - 1$
   \begin{gather*}
     \bar{X}_{v}(t) \equiv \hat{\pi}_1 \quad \text{for $v \leq m$} \\
     \bar{X}_{v}(t) \equiv \hat{\pi}_0 \quad \text{for $v > m$} 
   \end{gather*}
 \end{enumerate}
 where $\hat{\pi}_0$ and $\hat{\pi}_1$ are sub-probability vectors
 obtained by removing the last coordinate of $\pi_0$ and $\pi_1$. 
\end{definition}
The first approximation yields a sequence of independent random graphs
with independent edges. For $t \leq t^{*} - 1$, $G(t)$ is an
attributed instance of the Erd\"{o}s-Renyi random graphs. For $t \geq
t^{*}$, $G(t)$ is an attributed instance of the kidney-egg model
\cite{rukhin11} (see \figurename~\ref{fig:kidney-egg}), which in
itself is a special instance of the stochastic
block model \cite{Holland1983}. The unattributed kidney-egg model
can be denoted as $\kappa(n,m,p,q,s)$ with the quantities $p$, $q$,
and $s$ representing
\begin{gather*}
p = \mathbb{P}(u \sim v \, | \, u \leq m, v \leq m) = \langle
\hat{\pi}_1, \hat{\pi}_1 \rangle \\ q =
\mathbb{P}(u \sim v \, | \, u \leq m, v > m) = \langle \hat{\pi}_1,
\hat{\pi}_0 \rangle \\ s = \mathbb{P}(u
\sim v \, | \, u > m, v > m) = \langle \hat{\pi}_0, \hat{\pi}_0
\rangle . 
\end{gather*}
\begin{figure}[!tp]
  \centering
  \includegraphics[width=5cm]{graphics/Fig2-SSP2011-hat1.pdf}
  \caption{An illustration of the kidney-egg model for $G(t^{*})$. The
    probability of an edge is $\langle \hat{\pi}_0, \hat{\pi}_0
    \rangle$ for the connections inside the kidney set (black), $\langle
    \hat{\pi}_1, \hat{\pi}_1 \rangle$ for the connections inside the egg set
    (red), and $\langle \hat{\pi}_0, \hat{\pi}_1 \rangle$ for the connections
    between the kidney and egg set.}
  \label{fig:kidney-egg}
\end{figure}
 \begin{definition}
  \label{def:2}
  Let $\mathscr{V}$ be a rdpm with change-parameters $(t^{*}, m,
  \pi_0, \pi_1, \mathbf{Q}_0, \mathbf{Q}_1)$. Define
  $\mathbf{Z}_0$ and $\mathbf{Z}_1$ by
  \begin{gather}
    \mathbf{Z}_0 = (\mathbf{1}\mathbf{\pi}_0^{T} -
     \mathbf{Q}_0)^{-1}(\mathbf{I} - \mathbf{1}\pi_0^{T}) \\
    \mathbf{Z}_1 = (\mathbf{1}\mathbf{\pi}_1^{T} -
     \mathbf{Q}_1)^{-1}(\mathbf{I} - \mathbf{1}\pi_1^{T}).
  \end{gather}
  $\mathbf{Z}_0$ and $\mathbf{Z}_1$ are the fundamental matrices
  for the continuous-time Markov chain on $[K+1]$ with intensity
  matrix $\mathbf{Q}_0$ and $\mathbf{Q}_1$ (see
  e.g. \cite[p. 55]{asmussen03:_applied_probab_queues}). Let
  $\Sigma_0$ and $\Sigma_1$ be given by
  \begin{gather*}
    \Sigma_0 = \mathrm{diag}(\pi_0) \mathbf{Z}_0 +
    \mathbf{Z}_0^{T}
    \mathrm{diag}(\pi_0) \\
    \Sigma_1 = \mathrm{diag}(\pi_1) \mathbf{Z}_1 +
    \mathbf{Z}_1^{T}
    \mathrm{diag}(\pi_1).
  \end{gather*}
  The second order approximation $\widehat{\mathscr{V}}$ of
  $\mathscr{V}$ is the sequence $\{\widehat{X}(t),
  \widehat{G}(t)\}_{t=1}^{\infty}$ where
  \begin{enumerate}
  \item For each $t$ and each $v \in [n]$, $\widehat{X}_v(t)$ is a
    random vector obtained by truncating a multivariate normal random
    vector $Z_{v}(t)$ to $\mathscr{S}$ with mean and covariance
    matrices given below.
  \item For $t < t^{*} - 1$,
    \begin{equation}
      \mathbb{E}[Z_v(t)] \equiv \hat{\pi}_0, \,\, 
  \mathrm{Var}[Z_v(t)] \equiv \widehat{\Sigma}_0 \quad \text{for all
    $v$}.
    \end{equation}
  \item For $t \geq t^{*} - 1$,  
    \begin{gather*}
      \mathbb{E}[Z_v(t)] \equiv \hat{\pi}_1, \,\, 
\mathrm{Var}[Z_v(t)] \equiv \widehat{\Sigma}_{1} \quad \text{for $v
  \leq m$} \\
  \mathbb{E}[Z_v(t)] \equiv \hat{\pi}_0, \,\, \mathrm{Var}[Z_v(t)] \equiv \widehat{\Sigma}_{0} \quad \text{for $v
        > m$}
    \end{gather*}
    where $\widehat{\Sigma}_0$ and $\widehat{\Sigma}_1$ are the $K
    \times K$ matrices
    obtained by removing the last row and column of $\Sigma_0$ and
    $\Sigma_1$, respectively. 
  \end{enumerate}
\end{definition}
A second-order approximation yields a sequence of independent latent
position graphs
\cite{marchette08:_predic,scheinerman10:_model,hoff02:_laten}.

Suppose that we have a sequence $\{\mathscr{V}^{r} \colon r > 0 \}$ of
rdpm with vertices $[n]$ and attributes $[K]$ where for each $r > 0$,
$\mathscr{V}^{r}$ has change-parameters $(t^{*}, m, \pi_0, \pi_1, r
\mathbf{Q}_0, r \mathbf{Q}_1)$. The parameter $r$ can be thought of as
the vertex process rate, i.e., the waiting time decreases
exponentially as $r$ increases. Let us now consider the sequence of
first approximations $\bar{\mathscr{V}}^{r}$ and second approximations
$\widehat{\mathscr{V}}^{r}$ of $\mathscr{V}^{r}$. We note that
$\bar{\mathscr{V}}^{r_1} = \bar{\mathscr{V}}^{r_2}$ for any $r_1, r_2
> 0$. Let us then denote by $\bar{\mathscr{V}}$ the (a.e.) unique
first order approximation of $\mathscr{V}^{r}$ for $r > 0$. If we
denote by $\widehat{\Sigma}_{0}^{(r)}$ and
$\widehat{\Sigma}_{1}^{(r)}$ the matrices $\widehat{\Sigma}_0$ and
$\widehat{\Sigma}_1$ for $\widehat{\mathscr{V}}^{r}$, then
$\widehat{\Sigma}_0^{(r)} = \widehat{\Sigma}_0^{(1)}/r$ and
$\widehat{\Sigma}_{1}^{(r)} = \widehat{\Sigma}_1^{(1)}/r$. Therefore,
as $r \rightarrow \infty$, $\widehat{\Sigma}_0^{(r)} \rightarrow
\bm{0}$ and $\widehat{\Sigma}_1^{(r)} \rightarrow \bm{0}$ and so
$\widehat{\mathscr{V}}^{r} \overset{\mathrm{d}}{\longrightarrow}
\bar{\mathscr{V}}$. This indicates that for sufficiently large $r$,
there is little statistical difference among the
$\bar{\mathscr{V}}$, $\mathscr{V}^{r}$ and $\widehat{\mathscr{V}}^{r}$
(\cite[Theorem 2]{lee11}). The behavior of
$\widehat{\mathscr{V}}^{r}$ for ``small'' $r$ is less
clear. $\bar{\mathscr{V}}$ will thus serve as the basis for our
subsequent analysis on graph invariants for attributed random
graphs.

A comment should also be made about the utility of the exact
generative model. It is computationally intractable to estimate the
parameters of the underlying continuous time Markov chains, i.e. the
$\{\mathbf{Q}^{(v)}(t)\}$, given a snapshot of the graphs. Therefore,
the main interest of \cite{lee11}, and consequently of this paper,
lies not in the exact generative model but in its first and
second-order approximation. We will not be concerned with the
estimation of the latent position vectors for the first and
second-order approximation in this paper, but we note that they can be
done via spectral embedding techniques a la \cite{STFP-2011}. We also
note that more refined models, where the messaging events between each pairs of
vertices are modeled as point process,
e.g. \cite{Heard2010,PerryWolfe2010}, can also be formulated. The
snapshot of the graphs is then a binning of these point processes and
parameter estimation is fundamental to
these approaches.
\section{Change-point detection}
\label{sec:change-point-detect}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=8cm]{graphics/Fig1-SSP2011.pdf}
  \caption{Notional depiction of the problem of change-point detection
    in a time-series of graphs}
  \label{fig:notional_change_point}
\end{figure}
Let $\mathscr{V}$ be a random dot process model with change-parameters
$(t^{*}, m, \pi_0, \pi_1, \mathbf{Q}_0, \mathbf{Q}_1)$. The
change-parameters encapsulate a notion of chatter anomalies, i.e., a
subset of vertices of $\mathscr{V}$ with altered communication
behavior in an otherwise stationary setting as depicted in
Fig.~\ref{fig:notional_change_point}. We are interested in the problem
of testing, for a $t \in \mathbb{N}$, the hypotheses that $t$ is the
change-point of $\mathscr{V}$, namely
\begin{gather*}
  \mathscr{H}_0 \colon t^{*} > t \\
  \mathscr{H}_A \colon t^{*} = t
\end{gather*}
This will be done using the notion of fusion of attributed graph
invariants. The particular invariants of interest are the size
$\mathcal{E}$, number of triangles $\tau$, scan $\Psi$, and max degree
$\Delta$. The scan statistics $\Psi$ are introduced and applied to the
problem of detecting chatter anomalies in
\cite{priebe05:_scan_statis_enron_graph}. The number of triangles is
the simplest graph invariant based on clique counts and one can define
other test statistics for $k$-clique counts with $k \geq 4$.  The use
of various graph invariants as test statistics was considered in
\cite{pao11:_statis_infer_random_graph} and it was shown there, via
Monte Carlo analysis, that no single invariant is uniformly most
powerful.

Let $(G,\phi)$ be an attributed graph. For a given $\{u,v\} \in
\tbinom{V}{2}$, we define a $\Gamma_{uv} \in \mathbb{R}^{K}$ by
$\Gamma_{uv}(k) = \mathbb{I}\{\phi(\{u,v\}) = k\}$ where $\mathbb{I}\{
\cdot \}$ is the indicator function. Thus $\Gamma_{uv} = \bm{0} =
(0,0,\dots,0)$ unless $\{u,v\} \in E$. Under the independent edges
assumption, the $\{\Gamma_{uv}\}$ are independent. We consider linear
attribute fusion of graph invariants with parameter $\lambda \in
\mathbb{R}^{K}$ via
%\begin{gather}
%  \label{eq:6}
%  \mathcal{E}_{\lambda}(G) = \sum_{k=1}^{K} \lambda_k \sum_{u,v}
%  \mathbb{I}\{ \phi(uv) = k \} \\
%  \tau_{\lambda}(G) = \sum_{(i,j,k) \in K^{3}\vphantom{\tbinom{V}{3}}}
%  \, \, \sum_{(u,v,w)
%      \in \tbinom{V}{3}} \lambda_{i}\lambda_j \lambda_k h_{ijk}(u,v,w) \\
%  \Delta_{\lambda}(G) = \max_{v \in V} \, \sum_{k = 1}^{K} \lambda_k
%  \sum_{u \in N(v)}{\mathbb{I}\{\phi(\{u,v\}) = k\}} \\
%  \label{eq:3}
%  \Psi_{\lambda}(G) = \max_{v \in V} \sum_{k = 1}^{K}
%  \lambda_k \!\!\sum_{u,w \in N(v)}\!\!\! \mathbb{I}\{\phi(\{u,w\}) = k\} 
%  \end{gather}
\begin{align}
  \label{eq:6}
  \mathcal{E}_{\lambda}(G) &= \sum_{\mathclap{\{u,v\} \in
      \tbinom{V}{2}}} \langle
  \lambda, \Gamma_{uv} \rangle \\
  \tau_{\lambda}(G) &= \sum_{\mathclap{\{u,v,w\} \in \tbinom{V}{3}}} \langle
  \lambda, \Gamma_{uv} \rangle \langle \lambda,
  \Gamma_{uw} \rangle \langle \lambda, \Gamma_{vw}
  \rangle \\ 
\Delta_{\lambda}(G) &= \max_{v \in V} \, \sum_{w \in N(v)} \langle
\lambda, \Gamma_{vw} \rangle \\
\label{eq:3}
  \Psi_{\lambda}(G) &= \max_{v \in V} \sum_{u,w \in N[v]} \langle
  \lambda, \Gamma_{uw} \rangle.
  \end{align}
  where $N(v) = \{u \colon u \sim v\}$ is the set of neighbors of
  $v$ and $N[v] = v \cup N(v)$.

  Let $\{G(t)\}$ be a time series of graphs. Let $J_{\lambda}(t)$ be a
  statistic for $G(t)$ of the form as in Eq.~\eqref{eq:6} through
  Eq.~\eqref{eq:3}. We define the running average 
  $\bar{J}^{l}_\lambda(t)$ of $J_{\lambda}$ as
\begin{equation}
  \label{eq:4}
 \bar{J}^{l}_{\lambda}(t) = \frac{1}{l}\sum_{s = 1}^{l} J_{\lambda}(t - s) 
\end{equation}
where $l \in \mathbb{N}$ specifies the width of the running-average
window. Our main interest is in the normalized fusion statistic
$T_{\lambda}^{l}(t)$ as depicted in Fig.~\ref{fig:temporal}, namely
\begin{equation}
  \label{eq:7}
 T_{\lambda}^{l}(t) = % \begin{cases}
   % J_{\lambda}(t) - J_{\lambda}(t - 1) & \text{if $m = 1$} \\
   \frac{J_{\lambda}(t) -
     \bar{J}_{\lambda}^{l}(t)}{\sqrt{\tfrac{1}{l-1}
       \sum_{s=1}^{l}(J_{\lambda}(t - s) - \bar{J}_{\lambda}^{l}(t))^2}}
%   & \text{if $m \geq 2$}
 %  \end{cases}
\end{equation}
for $l \geq 2$. The choice of values for $l$ depends on the
practitioner, in that small values of $l$ can
induce large fluctuations in the temporal standardization statistics,
which may lead to a higher false positive rate while large values of $l$,
can induce over-smoothing, leading to a higher
false negative rate.

We are interested in finding the optimal fusion parameter $\lambda$,
i.e., we want to find a $\lambda$ such that
the power of the test using $T_{\lambda}^{l}$ is maximized. The
theoretical results in \S~\ref{sec:power-estimates} will be useful for
this purpose. We note that
$T_{\lambda}^{l}$ is scale invariant in $\lambda$ for all of our graph
invariants. It was noted in \cite{lee11} that the maximum of the power
for scale invariant test statistics is attainable in the set $\{ \lambda \colon \|
\lambda \| = 1 \}$. 
\begin{figure}[htbp]
  \centering
  \includegraphics[width=8cm]{graphics/Fig4-SSP2011.pdf}
  \caption{Temporal standardization: when testing for change at time
    $t$, the recent past (graphs $G(t - l), \dots, G(t-1))$ is used to
    standardize the invariants}
  \label{fig:temporal}
\end{figure}
\section{Power estimates}
\label{sec:power-estimates}
\S~\ref{sec:change-point-detect} described the framework of hypothesis
testing for anomaly/change-point detection in a time-series of
graphs. In particular, it was assumed that the graphs were generated
according to a random dot process model with change-parameters
$(t^{*}, m, \pi_0, \pi_1, \mathbf{Q}_0, \mathbf{Q}_1)$ and the test 
statistics were temporal standardization of several graph invariants, 
namely size, max degree, scan, and number of triangles.

We derive the limiting distributions for the above-mentioned test
statistics in this section. As we 
mentioned earlier, we will assume that the number of
vertices $n$ is large so that the use of the first order
approximation $\bar{\mathscr{V}}$ of $\mathscr{V}$ is appropriate for
studying the asymptotic theory for the above-mentioned graphs
invariants. We strive to obtain approximations for the power 
in testing $t^{*} > t$
against $t^{*} = t$ using $T_{\lambda}^{l}(t)$ and from these
approximations, find the $\lambda$ that maximize the power. 
The limiting distributions for graph
invariants in the {\em unattributed} setting are well-known, see 
e.g., \cite{bollobas85:_random_graph} for max-degree,
\cite{nowicki88:_subgr_u_statis_method} 
for subgraph counts, and \cite{rukhin12} for scan. The results
for the attributed and temporal standardization settings are thus 
straightforward extensions of
their counterparts in the unattributed settings.

We now introduce additional notation that will be used later on in the
paper. Let $\pi_{00}, \pi_{01}$, and $\pi_{11}$ be
sub-probability vectors whose components are given by
\begin{gather*}
  \pi_{00}(k) = \pi_{0}(k) \pi_{0}(k) \\
  \pi_{01}(k) = \pi_{0}(k) \pi_{1}(k) \\
  \pi_{11}(k) = \pi_{1}(k) \pi_{1}(k)
\end{gather*}
where $k \in [K]$. We also let $\eta_{00}, \eta_{01}$, and
$\eta_{11}$ be matrices of size $K \times K$ defined by 
\begin{gather*}
  \eta_{00} = \mathrm{diag}(\pi_{00}) - \pi_{00} \pi_{00}^{T} \\
  \eta_{01} = \mathrm{diag}(\pi_{01}) - \pi_{01} \pi_{01}^{T} \\
  \eta_{11} = \mathrm{diag}(\pi_{11}) - \pi_{11} \pi_{11}^{T}.
\end{gather*}

\subsection{Power estimates for $\mathcal{E}_\lambda$}
\label{sec:power-estim-mathc}
It was shown in \cite{lee11} that if $J_\lambda(t) =
\mathcal{E}_{\lambda}(G(t))$, then $T_{\lambda}^{l}(t)$ follows a
$t$-distribution with $l - 1$ degrees of freedom in the
limit. Specifically, we have the following results
\begin{theorem}
  \label{thm:9}
  Let $t \in \{l+1, \dots, t^{*}\}$ and $\lambda \in
  \mathbb{R}^{K}$. Define the vector $\zeta$ and the matrix $\xi$ by
  \begin{gather*}
    \zeta = \tbinom{m}{2}(\pi_{11} - \pi_{00}) + (n-m)m(\pi_{01} -
    \pi_{00}) \\
    \xi = \tfrac{l+1}{l}\tbinom{n}{2} \eta_{00} +
          \tbinom{m}{2}(\eta_{11} - \eta_{00}) + (n-m)m(\eta_{01} -
          \eta_{00}).
  \end{gather*}
  Suppose that $\xi$ is positive definite.
  Define the random variable $\psi^{l}_{\lambda}(t)$ by
  \begin{equation}
    \label{eq:10}
    \psi^{l}_{\lambda}(t) = \begin{cases}
      \sqrt{\frac{l}{l + 1}} T_{\lambda}^{l}(t)& \text{if $t < t^{*}$}
      \\ \sqrt{\frac{\langle \lambda, \tbinom{n}{2} \eta_{00}
            \lambda \rangle}{\langle \lambda, \xi
            \lambda \rangle}} T_{\lambda}^{l}(t) & \text{if $t =
          t^{*}$}.
      \end{cases}
  \end{equation}
As $n \rightarrow \infty$,
  $\psi^{l}_{\lambda}(t)$ converges weakly to the Student
  $t$-distribution with $l-1$ degrees of freedom and non-centrality
  parameter $\mu_{\lambda}$, where
  \begin{equation}
    \label{eq:15}
    \mu_{\lambda} = \begin{cases}
      0 & \text{if $t < t^{*}$} \\
      \frac{\langle \lambda, \zeta \rangle}{\sqrt{\langle \lambda, \xi \lambda \rangle}} & \text{if $t = t^{*}$} 
    \end{cases}
  \end{equation}
\end{theorem}
We are interested in finding the $\lambda$ that will maximize the
power of the test. The power approximation is determined by various
factors but the dominating factor is $\mu_{\lambda}$ (for large $n$). 
The following corollary extends a result in
\cite{lee11} for $K = 2$ to the case
where $K \geq 2$. 
\begin{corollary}
  \label{cor:1}
  Let $\zeta$ and $\xi$ be as defined in Theorem~\ref{thm:9}. Suppose
  that $\xi$ is also positive definite. Let $\nu^{*}$ be the
  normalized eigenvector corresponding to the largest eigenvalue of $ \xi^{-1/2}
  \zeta \zeta^{T} \xi^{-1/2}$, i.e.,
 \begin{equation}
   \label{eq:9}
  \nu^{*} = \argmax_{\nu \colon \| \nu \| = 1}
  \nu^{T} \xi^{-1/2} \zeta \zeta^{T} \xi^{-1/2}
  \nu.
 \end{equation}
 Then $\lambda^{*} = \tfrac{\xi^{-1/2} \nu^{*}}{\|\xi^{-1/2} \nu^{*} \|}$ satisfies
 \begin{gather}
   \label{eq:8}
 \argmax_{ \|\lambda\| = 1}
 \mu_{\lambda} \, \cap \, \{\lambda^{*}, - \lambda^{*}\} \not = \emptyset \\
 \argmin_{ \| \lambda \| = 1} \mu_{\lambda} \, \cap \, \{\lambda^{*}, -
 \lambda^{*}\} \not = \emptyset.
 \end{gather}
\end{corollary}
\subsection{Power estimates for $\tau_{\lambda}$}
\label{sec:power-estim-tau_l}
The limiting distribution for the number of triangles in unattributed
random graphs was considered in
\cite{nowicki88:_subgr_u_statis_method} for the Erd\"{o}s-Renyi and in
\cite{rukhin09:_asymp_analy_various_statis_random_graph_infer} for the
kidney-egg model. We note here the small changes that allow us to
extend the results in
\cite{rukhin09:_asymp_analy_various_statis_random_graph_infer,%
nowicki88:_subgr_u_statis_method} to our attributed graphs model. Let
$Y_{uv} = \langle \lambda, \Gamma_{uv} \rangle$ for $\{u,v\} \in
\tbinom{V}{2}$. We can now write $\tau_{\lambda}(G(t))$ as
\begin{equation}
  \label{eq:45}
  \frac{\tau_{\lambda}(G(t))}{\tbinom{n}{3}} = \tbinom{n}{3}^{-1}
  \sum_{\mathclap{\{u,v,w\} \in \tbinom{V}{3}}} Y_{uv} Y_{uw} Y_{vw};
\end{equation}
$\tau_{\lambda}(G(t))/\tbinom{n}{3}$ is then an
$U$-statistic on $\{Y_{e}\}$, with kernel function $h(Y_{1}, Y_{2},
Y_{3}) = Y_{1} Y_{2} Y_{3}$. By using Hajek's projection method, we
can show that $\tau_{\lambda}(G(t))$ converges to a normal
distribution as $n \rightarrow \infty$. 
\begin{lemma}
  \label{lem:3}
  Let $\tau_{\lambda}^{*}$ be the Hajek's projection of $\tau_{\lambda}$, i.e.,
\begin{equation}
  \label{eq:46}
  \tau_{\lambda}^{*}(G) - \mathbb{E}[\tau_{\lambda}(G)] =
  \sum_{\mathclap{\{u,v\} \in \tbinom{V}{2}}} \, \Bigl(\mathbb{E}[
  \tau_{\lambda}(G) \, \lvert \, Y_{uv}] -
  \mathbb{E}[\tau_{\lambda}(G)]\Bigr).
\end{equation}
For $t < t^{*}$, we have
\begin{align*}
  \mathbb{E}[\tau_{\lambda}^{*}(G(t))] &= \tbinom{n}{3} \langle 
  \lambda, \pi_{00} \rangle^{3} \\
  \sqrt{\mathrm{Var}[\tau_{\lambda}^{*}(G(t))]} &= (n-2) \langle \lambda,
  \pi_{00} \rangle^{2} \sqrt{\tbinom{n}{2} \langle \lambda, \eta_{00}
  \lambda \rangle}.
\end{align*}
For $t = t^{*}$, we have
  \begin{gather*}
    \label{eq:42}
    \begin{split}
    \mathbb{E}[\tau_{\lambda}^{*}(G(t))] = \tbinom{m}{3}
    \langle \lambda, \pi_{11} \rangle^{3} &+
    \tbinom{m}{2}(n-m) \langle \lambda, \pi_{11} \rangle \langle
    \lambda, \pi_{01} \rangle^{2} \\ &+ m \tbinom{n-m}{2} \langle \lambda,
    \pi_{01} \rangle^{2} \langle \lambda, \pi_{00} \rangle \\ &+
    \tbinom{n-m}{3} \langle \lambda, \pi_{00} \rangle^{3}
    \end{split} \\
    \begin{split}
    \mathrm{Var}[\tau_{\lambda}^{*}(G(t))] = \tbinom{m}{2} \langle \lambda, \eta_{11} \lambda
    \rangle S_1^2  
    &+ m(n-m) \langle \lambda, \eta_{01} \lambda \rangle S_2^2  \\
    &+ \tbinom{n-m}{2} \langle \lambda, \eta_{00} \lambda \rangle
    S_3^2
    \end{split}
  \end{gather*}
  where $S_1$, $S_2$, and $S_3$ are given by
\begin{align*}
    S_1 &= (m-2) \langle \lambda, \pi_{11} \rangle^{2} + (n-m)
    \langle \lambda, \pi_{01} \rangle^{2} \\ 
    S_2 &= (m-1)
    \langle \lambda, \pi_{11} \rangle \langle \lambda, \pi_{01}
    \rangle + (n-m-1) \langle \lambda, \pi_{00} \rangle \langle
    \lambda, \pi_{01} \rangle \\
    S_3 &= m \langle \lambda, \pi_{01} \rangle^{2} + (n-m-2) \langle
    \lambda, \pi_{00} \rangle^{2}.
    \end{align*}
  As $n \rightarrow \infty$, we have
  \begin{equation}
    \label{eq:49}
    \frac{\tau_{\lambda}(t) -
      \mathbb{E}[\tau_{\lambda}^{*}(G(t))]}{\sqrt{\mathrm{Var}[\tau_{\lambda}^{*}(G(t))]}}
\overset{\mathrm{d}}{\longrightarrow}  \mathcal{N}(0,1).
  \end{equation}
\end{lemma}
From Lemma~\ref{lem:3}, we can show that the limiting distribution of
$T_{\lambda}^{l}(t)$ is once again a Student t-distribution with
$l-1$ degrees of freedom.
\begin{theorem}
  \label{thm:5}
  Let $t \in \{l+1, \dots, t^{*}\}$ and $\lambda \in
  \mathbb{R}^{K}$. Let $\mu_0 = \mathbb{E}[\tau_{\lambda}^{*}(G(t))]$
  and $\sigma_0^{2} = \mathrm{Var}[\tau_{\lambda}^{*}(G(t))]$ for $t <
  t^{*}$. Let $\mu_A = \mathbb{E}[\tau_{\lambda}^{*}(G(t^{*}))]$ and
  $\sigma_{A}^{2} = \mathrm{Var}[\tau_{\lambda}^{*}(G(t^{*}))]$. Define
  the random variable $\psi_{\lambda}^{l}(t)$ by
  \begin{equation}
    \label{eq:50}
    \psi_{\lambda}^{l}(t) = \begin{cases}
      \sqrt{\tfrac{l}{l+1}}\, T_{\lambda}^{l}(t) & \text{if $t <
        t^{*}$} \\
      \sqrt{\tfrac{l \sigma_0^{2}}{l \sigma_A^{2} + \sigma_0^{2}} }\,
      T_{\lambda}^{l}(t) & \text{if $t = t^{*}$}.
        \end{cases}
  \end{equation}
  As $n \rightarrow \infty$, $\psi_{\lambda}^{l}(t)$ converges to a Student $t$-distribution with $l - 1$ degrees of freedom and
  non-centrality parameter $\mu_{\lambda}$ where
  \begin{equation}
    \label{eq:51}
    \mu_{\lambda} = \begin{cases} 0 & \text{if $t < t^{*}$} \\
      \frac{\mu_A - \mu_0}{\sqrt{l \sigma_A^{2} + \sigma_0^{2}}} &
       \text{if $t = t^{*}$}.
      \end{cases}
  \end{equation}
\end{theorem}
The power approximation for $\tau_{\lambda}$ is determined by various
factors, but similar to the power approximation for
$\mathcal{E}_{\lambda}$, the dominating factor is $\mu_\lambda$ (for
large $n$). Thus, we are also interested in finding the $\lambda$
that will maximize $\mu_\lambda$. However, because of the high power
of $\lambda$ in the expression for $\mu_{\lambda}$, an exact 
solution to $\argmax_{\lambda}{\mu_{\lambda}}$ is more challenging for
$\tau$ than for $\mathcal{E}$. 
\subsection{Power estimates for $\Delta_{\lambda}(t)$}
\label{sec:power-estim-delt}
The limiting distribution for maximum degree in unattributed random
graphs was considered in \cite{bollobas85:_random_graph} for the
Erd\"{o}s-Renyi and in \cite{rukhin11} for the kidney-egg model. We
note here the necessary changes that allow us to to extend the results
in \cite{bollobas85:_random_graph,rukhin11} to our attributed graphs
model. We denote by $\mathcal{G}(\alpha,\beta)$ the Gumbel
distribution with location parameter $\alpha$ and scale parameter
$\beta$.
\begin{proposition}
  \label{prop:3}
  Let $a_n$ and $b_n$ be functions of $n$ given by
  \begin{align*}
    a_n &= (2 \log{n})^{1/2}\Bigl(1 - \frac{\log{\log{n}} + \log{4\pi}}{4 \log{n}} \Bigr) \\ 
    b_n &= (2 \log{n})^{-1/2}.
  \end{align*}
  Then as $n \rightarrow \infty$ and $m = \Omega( \sqrt{n \log{n}})$,
  we have
  \begin{gather}
    \label{eq:14}
    \frac{\Delta_{\lambda}(t) - a_{n} \sigma_0 - \mu_0}{b_n \sigma_0}
   \overset{\mathrm{d}}{\longrightarrow}  \mathcal{G}(0,1) \quad \text{for $t < t^{*}$ } \\
    \frac{\Delta_{\lambda}(t) - a_{m} \sigma_A - \mu_A}{b_m \sigma_A}
       \overset{\mathrm{d}}{\longrightarrow}\mathcal{G}(0,1) \quad \text{for $t = t^{*}$ }
  \end{gather}
  where
  \begin{align*}
    \mu_0 &= (n-1)\langle \lambda, \pi_{00} \rangle \\
    \sigma_A &= \sqrt{(n-1)\langle \lambda, \eta_{00} \lambda \rangle} \\
    \mu_0 &= \,\, (m - 1) \langle \lambda, \pi_{11} \rangle + (n-
    m)\langle \lambda, \pi_{01} \rangle
    \\ \sigma_A &= \sqrt{ (m - 1) \langle \lambda, \eta_{11} \lambda \rangle + (n -
      m) \langle \lambda, \eta_{01} \lambda \rangle}.
    \end{align*}
\end{proposition}
$T_{\lambda}^{l}(t)$ based on the Gumbel distributed
$\Delta_{\lambda}(t)$, in contrast to $T_{\lambda}^{l}(t)$ based on
the normally distributed $\mathcal{E}_{\lambda}$ and
$\tau_{\lambda}(t)$, does not have a simple distribution for small or
moderate values of $l$. The following result gives the limiting
distribution for $T_{\lambda}^{l}(t)$ under the assumption that $l$ is
sufficiently large. 
\begin{theorem}
  \label{thm:8}
  Let $t \in \{l+1, \dots, t^{*}\}$ and $\lambda \in
  \mathbb{R}^{K}$. For sufficiently large $n$ and sufficiently large
  $l$, the variable $T_{\lambda}^{l}(t)$ has approximately a
  $\mathcal{G}(\rho_{\lambda}, \varsigma_{\lambda})$ distribution with
\begin{align}
  \label{eq:13}
  \rho_{\lambda} &= \begin{cases}
      - \frac{ \sqrt{6}}{\pi} \gamma & \text{if $t < t^{*}$} \\
    \frac{\sqrt{6}}{\pi} \frac{\mu_A - \mu_0 + a_m\sigma_A - (a_n + b_n
      \gamma)\sigma_0}{b_n \sigma_0} & \text{if $t =
        t^{*}$} 
  \end{cases}\\
  \varsigma_{\lambda} &= \begin{cases}
    \frac{\sqrt{6}}{\pi} & \text{if $t < t^{*}$} \\
    \frac{\sqrt{6}}{\pi}\frac{b_m \sigma_A}{ b_n \sigma_0} & \text{if $t =
      t^{*}$} \\
    \end{cases}
\end{align}
where $\gamma \approx 0.57721$ is the Euler-Mascheroni constant.
\end{theorem}
We are interested in finding the $\lambda$ that will maximize the
power of the test using $\Delta_{\lambda}$. The power approximation is
determined by various factors, but the dominating factor is
$\tfrac{\rho_\lambda}{\varsigma_\lambda}$. For sufficiently large $n$
and $m$, we have for $t = t^{*}$,
\begin{equation*}
  \begin{split}
  \frac{\rho_\lambda}{\varsigma_\lambda} &= (1 + o(1)) \frac{\mu_A - \mu_0}{b_m
    \sigma_A} \\ &= (1 + o(1)) \frac{\langle \lambda, (m-1)\pi_{11} + (n-m)\pi_{01} -
    (n-1) \pi_{00} \rangle}{b_m \sqrt{\langle \lambda, ((m-1) \eta_{11} +
    (n-m)\eta_{01}) \lambda \rangle}}.
  \end{split}
\end{equation*}
We can thus find the maximum and minimum of
$\tfrac{\rho_{\lambda}}{\varsigma_{\lambda}}$ by solving an 
eigenvalue problem as in Corollary~\ref{cor:1}.
\subsection{Power estimates for $\Psi_{\lambda}(t)$}
\label{sec:power-estim-psi_l}
The limiting distribution for the scan statistics in unattributed
random graphs was considered in
\cite{rukhin12}. We note here the changes
that allow us to extend the results in
\cite{rukhin12} to our attributed graphs
model. Let $z_n$ be defined, for $n \in \mathbb{N}$, by
\begin{equation}
  \label{eq:20}
   z_n = \sqrt{2
      \log{n}}\Bigl(1 - \frac{\log{\log{n}} + \log{4 \pi}}{4
      \log{n}}\Bigr).
\end{equation}
Let us also define two random variables $E$ and $F$ by
\begin{gather*}
 %  B \sim \mathrm{Bin}(m, \langle 1, \pi_{00} \rangle); \quad C \sim \mathrm{Bin}(n-m-1,
 % \langle 1, \pi_{01} \rangle) \\
  E \sim \mathrm{Bin}(m-1, \langle 1, \pi_{11} \rangle) \\ F \sim \mathrm{Bin}(n-m,
  \langle 1, \pi_{01} \rangle).
\end{gather*}%
Let $E + F$ be the convolution of $E$ and $F$. Denote by $\mu_{E+F}$ and
$\sigma_{E+F}^{2}$ the mean and variance of $E + F$. Let
$N_{\kappa} = \mu_{E + F} + z_m \sigma_{E + F}$. We then have the
following results.
\begin{lemma}
  \label{lem:5}
  Let $\Psi_{\lambda}(t)$ be the scan statistic for $t < t^{*}$. Let
  $N_{0}$, $a_n$, and $b_n$ be defined by
\begin{align*}
   \mu_0 &= (n-1) \langle 1, \pi_{00} \rangle \\
   \sigma_0 &= \sqrt{(n-1) (\langle 1, \pi_{00} \rangle - \langle
     1, \pi_{00} \rangle^2)} \\ 
    N_{0} &= \mu_0 + z_n \sigma_0 \\
  %  a_n &= N_{0} \frac{\langle \lambda, \pi_{00} \rangle}{\langle
  %    1, \pi_{00} \rangle} + C_{00}p_{00}\tbinom{N_{0}}{2} \\ 
    a_n &=  \langle \lambda, \pi_{00} \rangle \tbinom{N_{0}}{2} \\ 
   b_n &= \langle \lambda, \pi_{00} \rangle N_{0} \frac{\sigma_0}{\sqrt{2 \log
     n}}.
%   b_n &= (1 - \tfrac{C_{00}p_{00}}{2} + C_{00}p_{00} N_{0}) \frac{\sigma_0}{\sqrt{2 \log
%     n}}
\end{align*}
Then we have
\begin{equation}
  \frac{\Psi_{\lambda}(t) - a_{n}}{b_n}
  \overset{\mathrm{d}}{\longrightarrow}  \mathcal{G}(0, 1).
\end{equation}
\end{lemma}%
\begin{lemma}
  \label{lem:6}
  Let $a_{n,m}$ and $b_{n,m}$ be given by
  \begin{gather}
    \label{eq:28}
    \begin{split}
    a_{n,m} =%  N_{\kappa} \frac{\langle x, \pi_{01} \rangle}{\langle
 %      1, \pi_{01} \rangle} \\
 % & + 
\langle \lambda, \pi_{00} \rangle \tbinom{N_\kappa}{2} &+
\langle \lambda, \pi_{11} - \pi_{00} \rangle
\tbinom{\mu_E}{2} \\ &+ \langle \lambda, \pi_{01} - \pi_{00} \rangle \mu_E \mu_F 
    \end{split} \\
    b_{n,m} = \langle \lambda, \pi_{00} \rangle N_\kappa \frac{\sigma_{E + F}}{\sqrt{2
        \log{m}}}.
    %b_{n,m} = (1 - \tfrac{C_{00} p_{00}}{2} + C_{00} p_{00}
    %N_\kappa)\frac{\sigma_{E + F}}{\sqrt{2 \log{n_1}}}
  \end{gather}
  If $m = \Omega(\sqrt{n \log n})$ and $m = O(n^{k/(k+1)})$ for some
  $k \in \mathbb{N}$ then
  \begin{equation}
    \label{eq:29}
    \frac{\Psi_{\lambda}(t^{*}) - a_{n,m}}{b_{n,m}}
    \overset{\mathrm{d}}{\longrightarrow} \mathcal{G}(0,1).
  \end{equation}
\end{lemma}
The next result is analogous to Theorem \ref{thm:8} and gives 
the limiting distribution for
$T_{\lambda}^{l}(t)$ based on $\Psi_{\lambda}$ for the case of large
$l$.
\begin{theorem}
  \label{thm:6}
  Let $t \in \{l+1, \dots, t^{*}\} $ and $\lambda \in
  \mathbb{R}^{K}$. For sufficiently large $n$ and sufficiently large
  $l$, $T_{\lambda}^{l}(t)$ has approximately a
  $\mathcal{G}(\rho_{\lambda}, \varsigma_{l})$ distribution with
  \begin{align}
    \label{eq:52}
    \rho_{\lambda} &= \begin{cases}
      - \tfrac{\sqrt{6}}{\pi} \gamma & \text{if $t < t^{*}$} \\
      \tfrac{\sqrt{6}}{\pi} \tfrac{a_{n,m} - a_n - b_n \gamma}{b_n} & \text{if
        $t = t^{*}$} 
    \end{cases} \\
      \varsigma_{\lambda} &= \begin{cases}
        \tfrac{\sqrt{6}}{\pi} & \text{if $t < t^{*}$} \\
        \tfrac{\sqrt{6}}{\pi} \tfrac{b_{n,m}}{6 b_n} & \text{if $t = t^{*}$}. \\
      \end{cases}
  \end{align}
\end{theorem}
The dominating factor in the power approximation for $\Psi_{\lambda}$
is $\tfrac{\rho_{\lambda}}{\varsigma_{\lambda}}$. For $t = t^{*}$ and
sufficiently large $n$ and $l$, we have
\begin{equation}
  \label{eq:53}
  \begin{split}
    \frac{\rho_{\lambda}}{\varsigma_{l}} &= \frac{a_{n,m} - a_n - b_n
      \gamma}{b_{n,m}} \\
    &= (1 + o(1)) \frac{ \langle \lambda, \xi \rangle}{\langle
      \lambda, N_{\kappa} \tfrac{\sigma_{E+F}}{\sqrt{2 \log{n}}}
      \pi_{00} \rangle}
  \end{split}
\end{equation}
where $\xi$ is given by
\begin{equation*}
 \xi =  \Bigl(\tbinom{N_{\kappa}}{2} -
    \tbinom{N_0}{2} - \tbinom{\mu_E}{2} - \mu_E \mu_F\Bigr)\pi_{00} +
    \tbinom{\mu_{E}}{2} \pi_{11} + \mu_E \mu_F \pi_{10}.
\end{equation*}
From Eq.~\eqref{eq:53}, we see that there exists (as $n \rightarrow
\infty$) a $\lambda$ that maximizes
$\tfrac{\rho_\lambda}{\varsigma_{\lambda}}$ and is on the boundary,
i.e., $\lambda_k \not = 0$ for exactly one $k$. Therefore the asymptotic theory
for scan statistics, in contrast with the other graph invariants that
were considered, indicates that there may be no benefits in fusing
attributes. However, as \figurename~\ref{fig5:subfig_scan} shows, this
phenomenon does not hold in general for moderate values of
$n$. Similar observations about potential inaccuracies in using
asymptotic results to predict finite but large samples behavior
in testing using graph invariants were discussed in
\cite{rukhin11,priebe10:_you_i}.   
%\section{Experimental Results}
%\label{sec:experimental-results}
\section{Inference Examples}
\label{sec:inference-examples}
We present here two inference examples, one on simulated data and the
other on the Enron email data set. In the first inference example, we
simulated data from the model in \S~\ref{sec:latent-process-model}
with the following parameters:
\begin{equation*}
  K = 2, n = 100, m = 9, l = 10,
\end{equation*}
transition matrices
\begin{equation*}
  \label{eq:55}
  \begin{matrix}
  \mathbf{Q}_0 = \begin{bmatrix}
    -\tfrac{2}{3} & \tfrac{1}{6} & \tfrac{1}{2} \\
    1 & -1 & 0 \\
    -1 & 0 & -1 
  \end{bmatrix},
  & \mathbf{Q}_1 = \begin{bmatrix}
    -\tfrac{13}{7} & \tfrac{5}{7} & \tfrac{8}{7} \\
    1 & -1 & 0 \\
    1 & 0 & -1
  \end{bmatrix}
  \end{matrix},
\end{equation*}
and stationary probability vectors
\begin{equation*}
  \pi_0 = (0.10,0.30,0.60)^{T}, \quad \pi_1 = (0.25,0.40,0.35)^{T}.
\end{equation*}
The change-point $t^{*}$ for this example is $t^{*} = 11$. 
Power estimates for our attribute fusion statistics for this example
are presented in \figurename~\ref{fig:power-scale} and
\figurename~\ref{fig:power-estimate}. We consider the asymptotic,
first approximation, second approximation, and exact model power
estimates. \figurename~\ref{fig:power-scale} shows power as a function
of the vertex process parameter $r$ at a specific point, namely at
$\lambda^{*} = \argmax_{\lambda} \beta(\lambda)$.
\figurename~\ref{fig:power-estimate} shows power as a function of
angle $\theta$, where $\lambda = (cos(\theta), sin(\theta))$. 
Finally, \figurename~\ref{fig:qqplot-maxd} presents Q-Q plots
illustrating convergence of the temporal standardization statistics to
the limiting distributions given in \S~\ref{sec:power-estimates}.

A comment should be made regarding the application of the
theoretical results in \S~\ref{sec:power-estimates} to this simulation
example. The results in \S~\ref{sec:power-estimates} on the limiting
distributions of the temporal standardization statistics specified both
the type of distribution, i.e., Gumbel or Student's t-distribution,
along with the associated parameters, i.e., location and scale for the Gumbel
distribution and degrees of freedom and non-centrality parameter for
the Student's t-distribution. 
Empirical investigations demonstrate that, for the temporal standardization
statistics based on max degree, number of triangles, and scan, the limiting distributions
with the parameters as specified in Theorem~\ref{thm:5},
Theorem~\ref{thm:8} and Theorem~\ref{thm:6} are accurate large-sample
approximations for these statistics provided that the graphs have $n \geq 1000$ and $m \geq 50$.
The variance parameters as specified by the theoretical result
require correction for small $n$, as used in our experiment ($n = 100$).
For example, $\mathrm{Var}[\tau_{\lambda}^{*}(G(t))]$ in Lemma~\ref{lem:3} is a
good approximation of the variance term in the normal approximation
only if the ratio of $\mathrm{Var}[\tau_{\lambda}^{*}(G(t))] -
\mathrm{Var}[\tau_{\lambda}(G(t))]$, the difference in the variance
between the Hajek's projection and the true variance, to
$\mathrm{Var}[\tau_{\lambda}(G(t))]$ is small; this does not hold
for $n = 100$. Thus, the results for the limiting distributions for the above experiment are obtained by
correcting for the variance. In particular, we used the sample variance of the
graph invariants under the first-order approximation as the variance
for the limiting distribution.

% A comment should be made regarding the application of the
% theoretical results in \S~\ref{sec:power-estimates} to this simulation
% example. The results in \S~\ref{sec:power-estimates} on the limiting
% distribution of the temporal standardization statistics specified both
% the type of distribution, i.e., Gumbel or Student's t-distribution,
% along with the associated parameters, i.e., location and scale for the Gumbel
% distribution and degree of freedom and non-centrality parameter for
% the Student's t-distribution. According to empirical evaluations that
% are not reported in this paper, for the temporal standardization
% statistics based on max degree, number of triangles, and scan, the limiting distributions
% with the parameters as specified in Theorem~\ref{thm:5},
% Theorem~\ref{thm:8} and Theorem~\ref{thm:6} are accurate large-sample
% approximations for these statistics provided that the graphs have $n
% \geq 1000$ and $m \geq 50$ and not for $n = 100$ as used in the above
% experiment. This is due mostly to the fact that the variance parameters as
% specified by the theoretical result is
% inaccurate for these small values of $n$. For example,
% $\mathrm{Var}[\tau_{\lambda}^{*}(G(t))]$ in Lemma~\ref{lem:3} is a
% good approximation of the variance term in the normal approximation
% only if the ratio of $\mathrm{Var}[\tau_{\lambda}^{*}(G(t))] -
% \mathrm{Var}[\tau_{\lambda}(G(t))]$, the difference in the variance
% between the Hajek's projection and the true variance, to
% $\mathrm{Var}[\tau_{\lambda}(G(t))]$ is small and this does not hold
% for $n = 100$. Thus, the results for the limiting distributions for the above experiment are obtained by
% correcting (and only correcting) for the variance. In particular, we used the sample variance of the
% graph invariants under the first-order approximation as the variance
% for the limiting distribution.

The main implication that can
be inferred from \figurename~\ref{fig:power-scale}, \figurename~\ref{fig:qqplot-maxd} and
\figurename~\ref{fig:power-estimate}
is that inferential performance in the
mathematically tractable first-order and second-order approximation
models does provide guidance for methodological choices applicable to
the exact (realistic but intractable) model. Furthermore, to the
extent that the exact model is realistic, we may tentatively conclude
that approximation model investigations have some bearing on real
data applications.
\begin{figure*}[!t]
  \centering
  \subfloat[$\mathcal{E}_{\lambda}$]{
    \includegraphics[width=7cm]{graphics/pwr-scale-size.pdf}
  }
  \hfil
  \subfloat[$\Delta_{\lambda}$]{
    \includegraphics[width=7cm]{graphics/pwr-scale-maxd.pdf}
  }
  \hfil
  \subfloat[$\tau_{\lambda}$]{
    \includegraphics[width=7cm]{graphics/pwr-scale-scan.pdf}
  }
  \hfil
  \subfloat[$\Psi_{\lambda}$]{
    \includegraphics[width=7cm]{graphics/pwr-scale-ntri.pdf}
  }
  \caption{Power $\beta(\theta^{*}_r,r)$ at $\theta_{r}^{*} =
    \argmax_{\theta} \beta(\theta,r)$ as a function of $r$ for the
    four invariants at test size $\alpha = 0.05$. The horizontal lines
    represents first-order approximation $\pm$ three standard
    deviations, and the two curves represent the second approximation
    (green) and exact model (blue). The 10000 Monte Carlo replicates
    yield standard deviations not exceeding $0.005$ for the power
    estimates . The second approximation results match well with the
    exact model results, and both match well for large $r$ with the
    first-order approximation results.  }
  \label{fig:power-scale}
\end{figure*}

\begin{figure*}[!t]
  \centering
  \subfloat[$T_{\Delta}$ under $H_0$]{
  \includegraphics[width=5cm]{graphics/maxd_null.pdf}
  }
  \hfil
  \subfloat[$T_{\tau}$ under $H_0$]{
  \includegraphics[width=5cm]{graphics/numtri_null.pdf}
  }
  \hfil
  \subfloat[$T_{\psi}$ under $H_0$]{
  \includegraphics[width=5cm]{graphics/scan_null.pdf}
  }
  \hfil
  \subfloat[$T_{\Delta}$ under $H_A$]{
  \includegraphics[width=5cm]{graphics/maxd_alt.pdf}
  }
  \hfil
  \subfloat[$T_{\tau}$ under $H_A$]{
  \includegraphics[width=5cm]{graphics/numtri_alt.pdf}
  }
  \hfil
  \subfloat[$T_{\psi}$ under $H_A$]{
  \includegraphics[width=5cm]{graphics/scan_alt.pdf}
  }
  \caption{Q-Q plots of the temporal standardization statistics
    $T_{\Delta}$, $T_{\tau}$ and $T_{\Psi}$ against the corresponding
    limiting distributions as given in Theorem~\ref{thm:5},
    Theorem~\ref{thm:8} and Theorem~\ref{thm:6}, respectively. The
    sample quantile for the temporal standardization statistics are
    based on 10000 Monte Carlo replicates of the time-series of
    graphs. The graphs are generated according to the second-order
    approximation with $r = 1024$. Refer to the text at the beginning
    of \S~\ref{sec:inference-examples} for details on the parameters
    of the random graph model. The fusion parameter $\lambda =
    (cos(\theta), sin(\theta))$ for each temporal standardization
    statistics correspond to the $\theta^{*}$ that maximizes the power
    as depicted in \figurename~\ref{fig:power-estimate}, i.e.,
    $\theta^{*}_{\Delta} \approx 0.32$, $\theta^{*}_{\tau} \approx 0.3$,
    $\theta^{*}_{\psi} \approx 0.14$.  For each plot, a linear regression
    between the sample quantile and the theoretical quantile was
    performed and the resulting best linear fit is plotted. The $R^2$
    statistics for the linear regressions exceed $0.995$ in all
    cases. }
  \label{fig:qqplot-maxd}
\end{figure*}

\begin{figure*}[!t]
  \centering
  \subfloat[$\mathcal{E}_{\lambda}$]{
    \includegraphics[width=7cm]{graphics/pwr-size.pdf}
  }
  \hfil
  \subfloat[$\Delta_{\lambda}$]{
    \includegraphics[width=7cm]{graphics/pwr-maxd-new.pdf}
  }
  \hfil
  \subfloat[$\Psi_{\lambda}$]{
    \includegraphics[width=7cm]{graphics/pwr-scan-new.pdf}
    \label{fig5:subfig_scan}
  }
  \hfil
  \subfloat[$\tau_{\lambda}$]{
    \includegraphics[width=7cm]{graphics/pwr-ntri-new.pdf}
  }
  \caption{Power $\beta$ as a function of angle $\theta$ for $\lambda
    = (cos(\theta),sin(\theta))$. The plot shows analytic asymptotic
    result (black), and first approximation (red), second
    approximation (green), and exact model (blue) estimates for the
    various invariants via Monte Carlo. The multiple green and blue
    lines correspond to different values of the vertex process rate
    $r$ (here $r \in \{2^{0}, 2^{1}, \dots, 2^{10}\}$). 
    The 10000 Monte Carlo
    replicates yields standard deviations not exceeding $0.005$ for
    the power estimates. The four vertical lines in each plot
    correspond to $\theta \in \{0, \pi/2, \pi/4,
    \theta^{*}\}$. Because we are considering a one-sided test and
    because $\pi_{1,k} > \pi_{0,k}$ for both $k = 1$ and $k = 2$, the
    power is maximized in the first quadrant, i.e., $\theta \in (0,
    \pi/2)$. The second approximation results match well
    with the exact model results, and both match well for large $r$
    with the first order approximation and the asymptotic results. The
    optimal $\theta^{*}$ is apparently different for the four different graph
    invariants.  
  }
  \label{fig:power-estimate}
\end{figure*}

Our second example uses the Enron data set which consists of email
messages among 184 executives of the Enron corporation during a time
period from 1998 to 2002. The Enron email messages for most of the
2001 calendar year have been annotated into 32 different topics 
\cite{berry01:_topic_annot_enron_email_data_set} and thus form the
basis of our analysis. We chose to condense these 32 different
topics into two groups. The first group consists of
general energy-related topics and the second group consists of
topics that are specifically related to the Enron corporation. 
From the collection of email messages, we construct a time series of
graphs. Each graph consists of email messages sent during a week in
2001. The vertices of the graphs represent the executives and the edges
represent email communication between the executives. The attributes on
the edges are given by the grouping of the email topics as mentioned
above.

\begin{figure}[tbp]
  \centering
 \includegraphics[width=8cm]{graphics/enron152.pdf} 
 \caption{Detecting chatter anomaly in the Enron time series of
   graphs. The plot shows the normalized statistics
   $T_{\lambda}^{l}(t)$ as a function of the fusion parameter $\lambda
   = (\cos(\theta), sin(\theta))$ for $l = 20$ and $t = 152$. The
   graph in question ($t = 152$) corresponds to the first week of May
   2001. The dashed lines correspond to the $\theta^{*}$ for the
   different graph invariants ($\theta^{*}_{\mathcal{E}_{\lambda}}
   \approx \theta^{*}_{\Delta_{\lambda}} \approx 0.96$,
   $\theta^{*}_{\Psi_{\lambda}} \approx 1.06$ and
   $\theta^{*}_{\tau_{\lambda}} \approx 0.83$). Attribute fusion
   provides superior detection.}
  \label{fig:enron}
\end{figure}

\begin{figure}[tbp]
  \centering
 \includegraphics[width=8cm]{graphics/Tstat_thetastar.pdf} 
 \caption{The fluctuations of $T_{\lambda}^{l}(t)$ over time for the
   four graph invariants. The value of $\lambda = (\cos(\theta),
   sin(\theta))$ for each graph invariant correspond to the optimal
   fusion parameter for the detection as illustrated in
   \figurename~\ref{fig:enron} and they are
   $\theta^{*}_{\mathcal{E}_{\lambda}} \approx
   \theta^{*}_{\Delta_{\lambda}} \approx 0.96$,
   $\theta^{*}_{\Psi_{\lambda}} \approx 1.06$ and
   $\theta^{*}_{\tau_{\lambda}} \approx 0.83$. The fluctuations
   indicate that there are possible anomalies around the weeks 120,
   132, and 152, which were also reported and analyzed in 
   \cite{priebe05:_scan_statis_enron_graph}. }
\label{fig:enron_time}
\end{figure}
The detection of a chatter anomaly in the Enron data set during the
first week of May 2001 ($t = 152$) was previously
reported in \cite{priebe05:_scan_statis_enron_graph}, but with 
unattributed edges. We chose to investigate the effect
of attribute fusion for that same week in our inference example. 
\figurename~\ref{fig:enron} presents a plot of the normalized
statistic $T_{\lambda}^{l}(t)$ at $t = 152$ with $l = 20$ for the four graph
invariants against the fusion parameter $\lambda = (cos(\theta),
sin(\theta))$. By using the limiting results of
\S~\ref{sec:power-estimates}, approximate p-values for hypothesis testing using these
$T_{\lambda}^{l}$ statistics can also be obtained. Under
the null hypothesis of homogeneity or no chatter groups, the
large-sample results of \S~\ref{sec:power-estimates} suggest that
$T_{\lambda}^{l}$ based on size and number of triangle graph
invariants follows a Student's t-distribution with $l - 1$ degrees of
freedom, while $T_{\lambda}^{l}$ based on max degree and scan graph
invariants follows a Gumbel distribution with location parameter
$-\tfrac{\sqrt{6}}{\pi} \gamma$ and scale parameter
$\tfrac{\sqrt{6}}{\pi}$. Thus we obtain p-values of
$0.01$, $0.002$, $0.003$, and $0.01$ for
$T_{\lambda}^{l}$ based on size, max degree, number of triangles and
scan, at $\theta^{*}_{\mathcal{E}_{\lambda}},
\theta^{*}_{\Delta_{\lambda}}, \theta^{*}_{\tau_{\lambda}},
\theta^{*}_{\Psi_{\lambda}}$, respectively. Ignoring the attributes on
the edges gave p-values of $0.13$, $0.02$, $0.51$ and $0.04$ for
$T_{\lambda}^{l}$ based on size, max degree, number of triangles, and
scan respectively. Finally, ignoring the edges with edge attribute 1
gave $p$-values of $0.02$, $0.004$, $0.005$, and $0.01$. Thus, in
general, attribute fusion provides superior detection as 
exemplified by the fact that the maximum for each of the normalized
statistics $T_{\lambda}$, and hence the minimum of the p-values, 
occurs for $\theta \not \in \{0,\pi/2\}$. Furthermore, the optimal 
fusion parameters are dependent on the specific graph invariants.

\figurename~\ref{fig:enron_time} presents a plot of the
normalized statistics $T_{\lambda}^{l}(t)$ with $l = 20$ for the four
graph invariants over the time interval corresponding to the
weeks $t = 115$ through $t = 187$. The value of $\lambda$ used
for each graph invariant is the $\lambda$ that maximizes
$T_{\lambda}^{l}(152)$. The fluctuations
indicate that there are possible anomalies around the weeks 120,
132, and 152, which were also reported and analyzed in 
\cite{priebe05:_scan_statis_enron_graph}.  Approximate p-values for
hypothesis testing using these $T_{\lambda}^{l}$ are
also available. For example, at week $t = 129$, the p-values are
$\approx 7 \times 10^{-4}, 4 \times 10^{-5}, 4 \times 10^{-5}, 2
\times 10^{-6}$ for $T_{\lambda}^{l}$ based on the size, max degree, number of
triangles, and scan statistic, respectively. At week $t = 125$, the
approximate p-values are $1 \times 10^{-5}, 9 \times 10^{-3}, 1
\times 10^{-6}$, and $8 \times10^{-4}$.
\section{Conclusions}
We have presented an analysis of change-point detection for time
series of attributed graphs through the use of test statistics which
are based on linear attribute fusion of some graph invariants. 
We derived the limiting distribution of these test statistics under
the assumptions that $n$, the number of vertices, is sufficiently
large and $r$, the process parameter rate, is also sufficiently
large. The limiting distribution for these test statistics are then
used to derive estimates for the power of the tests.

The simulation experiment in \S~\ref{eq:55} indicates that the power estimates
are accurate, even for the moderate value of $n = 100$. Furthermore,
it was also indicated that the optimal linear fusion parameter depends
on the graph invariant considered. In particular, the results depicted
in \figurename~\ref{fig:power-estimate} yield
$\theta^{*}_{\mathcal{E}_{\lambda}} \approx 0.24$,
$\theta^{*}_{\Delta_{\lambda}} \approx 0.32$,
$\theta^{*}_{\Psi_{\lambda}} \approx 0.14$ and
$\theta^{*}_{\tau_{\lambda}} \approx 0.38$. These optimal fusion
parameter differences are statistically significant, and combining
this result with the ``no uniformly most powerful invariant'' result
\cite{pao11:_statis_infer_random_graph}, we conclude that optimal
linear attribute fusion theory requires significant additional
development. Toward this end, the approximation models from
\cite{lee11} promise to be of assistance.

The principal utility of the large-approximation
results of \S~\ref{sec:power-estimates} is to the specification of
approximate p-values and the investigation on the Enron
corpus data set illustrates this succinctly.

% The exact generative model for time-series
% of attributed graphs as presented in \S~\ref{sec:power-estimates}
% contains assumptions that are not readily verified. However, as our
% analysis

Hypothesis testing on time series of attributed graphs has
applications in diverse areas, e.g., social network analysis (wherein
vertices represent individual actors or organizations), connectome
inference (wherein vertices are neurons or brain regions) and text
processing (wherein vertices represent authors or documents). These
and many other applications may benefit from generalizations of our
results and the model in \cite{lee11} to directed, multi, and weighted
graphs, to other class of graph invariants, and to 
consideration of inference with errorful edge
attributes through an attribute confusion matrix. For example, the
paper considered a test statistic based on the number of triangles
which is the simplest clique counts statistic. One can define
attribute fusion for $k$-clique counts for $k \geq 4$. Using the
theory of $U$-statistics for subgraphs counts
(cf. \cite{nowicki88:_subgr_u_statis_method}), the limiting
distribution for the temporal standardization statistics based on 
these $k$-clique counts can be derived in an analogous manner to the 
case for $k = 3$ in this paper.

Finally, we remark that applicability of the statistics
described in \S~\ref{sec:change-point-detect} extends beyond the
statistical inference perspective. Indeed, temporal standardization of
graph invariants can be computed irrespective of any underlying model
on the graphs. As such, these statistics can
serve as the basis for simple and robust inference procedures on
time-series of (attributed) graphs. 
\section*{Acknowledgements}
The authors thank the anonymous reviewers for their comments which
have improved the presentation and quality of the paper. This work was 
partially supported by National Security Science and Engineering
Faculty Fellowship (NSSEFF) and by Johns Hopkins University Human
Language Technology Center of Excellence (JHU HLT COE).
\bibliography{ssp2011}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./minh.jpg}}]{Minh Tang}
  received the BS degree from Assumption University,
  Thailand in 2001, the MS degree from the University of
  Wisconsin-Milwaukee in 2004 and the Ph.D degree from Indiana
  University Bloomington in 2010. He is currently a postdoctoral
  fellow in the Department of Applied Mathematics and Statistics at
  Johns Hopkins University. His research interests include statistical
  pattern recognition, dimensionality reduction, and high-dimensional
  data analysis.
  \end{IEEEbiography}

  \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./youngser.jpg}}]{Youngser Park}
    received the B.E. degree in electrical engineering from Inha
    University in Seoul, Korea in 1985, the M.S. and Ph.D. degrees in
    computer science from The George Washington University in 1991 and
    2011 respectively. From 1998 to 2000 he worked at the Johns
    Hopkins Medical Institutes as a senior research engineer. From
    2003 until 2011 he worked as a senior research analyst, and has
    been an associated research scientist since 2011 in the Center for
    Imaging Science at the Johns Hopkins University. At Johns Hopkins,
    he holds joint appointments in the Department of Applied
    Mathematics and Statistics and the Human Language Technology
    Center of Excellence. He has reviewed papers for ACM Transactions
    on Knowledge Discovery in Data, Statistical Analysis and Data
    Mining, and WIREs Computational Statistics. His current research
    interests are clustering algorithm, pattern classification, and
    data mining for high-dimensional and graph data.
\end{IEEEbiography}

  \begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./nam.jpg}}]{Nam H. Lee}
    Nam Lee received the Ph.D degree in 2008 from University of California ,
San Diego, from which he also received the MS degree in 2004 and the
BA degree in 2002. He is currently a faculty member of Department of Applied Mathematics
and Statistics at Johns Hopkins University.  His research interests
include stochastic processes and their applications to statistical pattern
recognition and optimal control.
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./cep.jpg}}]{Carey E. Priebe}
  received the B.S. degree in
  mathematics from Purdue University in 1984, the M.S. degree in
  computer science from San Diego State University in 1988, and the
  Ph.D. degree in information technology (computational statistics)
  from George Mason University in 1993. From 1985 to 1994 he worked as
  a mathematician and scientist in the US Navy research and
  development laboratory system. Since 1994 he has been a professor in
  the Department of Applied Mathematics and Statistics, Whiting School
  of Engineering, Johns Hopkins University, Baltimore, Mary- land. At
  Johns Hopkins, he holds joint appointments in the Department of
  Computer Science, the Department of Electrical and Computer
  Engineering, the Center for Imaging Science, the Human Language
  Technology Center of Excellence, and the Whitaker Biomedical
  Engineering Institute. He is a past President of the Interface
  Foundation of North America - Computing Science \& Statistics, a past
  Chair of the American Statistical Association Section on Statistical
  Computing, a past Vice President of the International Association
  for Statistical Computing, and on the editorial boards of Journal of
  Computational and Graphical Statistics, Computational Statistics and
  Data Analysis, and Computational Statistics. His research interests
  include computational statistics, kernel and mixture estimates,
  statistical pattern recognition, statistical image analysis,
  dimensionality reduction, model selection, and statistical inference
  for high-dimensional and graph data.  He is a Senior Member of the
  IEEE, a Lifetime Member of the Institute of Mathematical Statistics,
  an Elected Member of the International Statistical Institute, and a
  Fellow of the American Statistical Association.

\end{IEEEbiography}

\end{document}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: